### 2010-12-12 opencl/papers/foropencl.tex
a bit of history

%
% I'm adding an outline here so I can see how everything fits together

%
% \title{Exploiting First-Class Arrays in Fortran for Accelerator Programming}
%

%
% \abstract
%

%
% \section{Introduction}
%    \subsection{Approach}
%    \subsection{Why Fortran?}
%    \subsection{Comparison to Other Languages}
%

%
% \section{Programming model}
%    \subsection{Fortran Syntax}   
%       \subsubsection*{Array notation}
%       \subsubsection*{Elemental functions}
%       \subsubsection*{Pure procedures}
%       \subsubsection*{Shift functions}
%       \subsubsection*{Regions}
%    \subsection{New Functions}
%    \subsection{Parallelism}
%    \subsection{Limitations}
%

%
% \section{Source-To-Source Transformations}
%    \subsection{ForOpenCL}
%       \subsubsection{array syntax}
%       \subsubsection{where construct}
%    \subsection{New functions}
%       \subsubsection{region}
%    \subsection{Static Analysis}
%       \subsubsection{Analysis not required}
%    \subsection{Simplifying Assumptions}
%

%
% \section{Shallow Water Model}
%    \subsection{Equations}
%

%
% \section{Performance}
%

%
% \section{Conclusions}
%

\documentclass[10pt, conference, compsocconf]{IEEEtran}

\usepackage{cite}
\usepackage{graphicx}
\usepackage[cmex10]{amsmath}
\usepackage{array}
\usepackage{url}

\title{Exploiting First-Class Arrays in Fortran for Accelerator Programming}

%%
%% reorder appropriately later
%%
\author{\IEEEauthorblockN{Matthew J. Sottile}
\IEEEauthorblockA{Galois, Inc.\\
421 SW 6th Ave. Suite 300 \\
Portland, OR 97204\\
Email: matt@galois.com}
\and
\IEEEauthorblockN{Craig E Rasmussen}
\IEEEauthorblockA{Los Alamos National Laboratory\\
CCS-7, MS B287\\
Los Alamos, NM\\
Email: rasmussn@lanl.gov}
}

\begin{document}

\maketitle

\begin{abstract}
Emerging architectures for high performance computing often are well
suited to a data parallel programming model.  This paper presents a
simple programming methodology based on existing languages and compiler
tools that allows programmers to take advantage of these systems.
We will work with the array features of Fortran 90 to show how this
infrequently exploited, standardized language feature is easily
transformed to lower level accelerator code.  Our transformations are
based on a mapping from Fortran 90 to C++ code with OpenCL extensions.
\end{abstract}

\input{intro}
\input{programming-model}
\input{source-to-source}
\input{static-analysis}

\section{Performance}

\section{Conclusions}

%% Explain what we have accomplished (copied from proposal)

%% NEW - CER
The sheer complexity of programming for clusters of many or multi-core
processors with tens of millions threads of execution make the simplicity of
the data parallel model attractive.  Furthermore, the increasing complexity of
todays applications (especially when convolved with the increasing complexity
of the hardware) and the need for portability across hardware architectures
make a higher-level and simpler programming model like data parallel
attractive.

%% NEW - CER
The goal of this work has been to exploit source-to-source transformations that
allow programmers to develop and maintain programs at a high-level of
abstraction, without coding to a specific hardware architecture.
Furthermore these transformations allow multiple hardware architectures
to be targeted without changing the high-level source.  It also removes the
necessity for application programmers to understand details of the accelerator
architecture or to know OpenCL.

% Summarize results

\section{Junk}

\cite{chamberlain04zpl, roth97stencils}

\bibliographystyle{IEEEtran}
\bibliography{IEEEabrv,foropencl}

\end{document}