### 2010-12-14 opencl/papers/intro.tex
text massaging

\section{Introduction}
\label{sec:intro}

This paper presents a compiler-level approach for targeting a single
program to multiple fundamentally different low level execution and
programming models while allowing the application programmer to adopt
a single high level programming model.  We show that features of
Fortran 90 for data parallel programming are well suited to automatic
transformation to generate code specifically tuned for different
low-level programming models such as OpenCL and CUDA.  For algorithms
that can be easily expressed in terms of whole array, data parallel
operations, writing code in Fortran and transforming it automatically
to specific low level implementations removes the burden from the
programmer of working with tedious, error prone, low-level tools.

In the ideal situation, application programmers would like to adopt
a programming model in which they write their application once and
use automated tools to retarget it to many architectures.  This has
proven to be very challenging historically due to the subtle balance
between high-level expressiveness of code and the performance of the
lower-level code that is emitted by a compiler.  This ideal high-level
model that programmers work with should emphasize readability,
maintainability, and close proximity in abstraction to the problem being
solved.  It should not be corrupted with details of specific target
architectures solely for the purpose of single-system performance.
For certain classes of applications, specifically those that map well
onto a data-parallel programming model, we show that Fortran 90 contains
language features that encourage high-level programming abstractions
without sacrificing performance during low-level code generation.

%1. Goal is to write once, transform many. Otherwise potentially reprogram for every architecture.
%2. Goal is to code for readability and maintainability, not performance
%3. This goal requires expression in a high-level language.
%4. But language must be simple enough for compiler to analyze.
%5. Thus ideal if language maps well to accelerator architectures.
%6. Data parallel constructs in Fortran 90 are chosen.
%7. Up to 65 times speed up measured on automatically transformed code.

At Los Alamos National Laboratory (LANL), as with many supercomputing
facilities today, users have a wide variety of computer platforms from
which to choose.  The most common platform is made up of clusters of
compute nodes with standard multi-core processors.  An increasingly
common feature is that some nodes also have accelerators that range
from the IBM Cell processor (such as the LANL Roadrunner system) to a
variety of GPUs from NVIDIA and AMD.  Some nodes have hardware with
vector instructions and others do not.  The peak performance of these
accelerated nodes often resides in the hundreds of gigaflops.

%The peak performance of the
%accelerated nodes range from 200+ GFlops for the IBM Cell/B.E.
%processor to XXX for NVIDIA Fermi. %% Fermi has another name

However, the performance that these accelerator architectures offer
comes at a cost as processor architectures are trending toward
multiple cores with instances of integrated accelerator units with
user managed memory and less of a reliance on superscalar instruction
level parallelism and hardware managed memory hierarchies (such as
traditional caches).  These changes place a heavy burden on
application programmers as they are forced to adapt to the new
systems.  An especially challenging problem faced by application
programmers is not only how to program to these new architectures
(considering the massive scale of concurrency available), but also how
to design programs that are portable across the changing landscape of
computer architectures with unique memory systems and programming
models.  The fundamental question we address is what programming model
and language constructs are best suited to span this set of new
hardware designs.

%% NEW - CER
%%%We will examine how the existing data-parallel constructs in Fortran can be combined with coarrays or MPI to provide effectively a new parallel programming language, one that is evolutionary in nature and provides complete compatibility with existing applications and libraries.  Data parallelism is a high-level abstraction that is, at the same time, both easier to program and gives the compiler more leeway (if fully exploited) in retargeting a program to different computer architectures.

A common theme amongst new processors is the emphasis on data parallel
programming.  This model is well suited to emerging architectures that
are based on either vector processing or massively parallel
collections of simple cores.  A popular trend for programming systems
such as GPU accelerators is the CUDA language from NVidia, or new
languages like Chapel from Cray.  A less intrusive approach that is
well suited to legacy applications and languages are language
extensions or libraries that allow programmers to avoid adopting
entirely new languages.  The recent OpenCL specification is intended
to support this programming model, as are directive-based methods such
as OpenMP or the Accelerator programming model from the Portland
Group~\cite{pgi10accelerator}.

The problem with many of these choices is that they expose too much
detail about the machine architecture to the programmer.  This is
particularly true of CUDA and OpenCL.  In CUDA, programmers must adapt
their codes to fit the threading model used by NVidia GPUs, while
OpenCL requires programmers to provide specially tuned versions of
their code for different classes of machine.  In both cases, the
programmer is also responsible for explicitly managing memory,
including staging of data back and forth from the host CPU and the
accelerator device memory.  While these models have been attractive as
a method for early adopters to utilize these new architectures, they
are less attractive to programmers who do not have the time or
resources to manually port their code to every new architecture and
programming model that emerges.

At this point in time it is not really possible to write once and run
efficiently on the wide variety of computer platforms we have
available.  For some classes of applications, we believe that this
goal is possible using language constructs already present in a
popular mainstream scientific programming language -- Fortran.  A
common, long-standing tongue-in-cheek response to new language
developments in the scientific and high performance computing
community is that a new language will arise to answer the needs of new
systems, and it will be called Fortran.  We believe that the work
presented in this paper validates that notion -- we need a new
language to work with, and that language is Fortran.

\subsection{Approach}

This paper addresses the accelerator programmign problem by examining
features in Fortran that allow programmers to express algorithms at a
very high level that can be easily transformed by a compiler to run
efficiently on a wide variety of platforms.  In particular we consider
computers based on GPUs and related accelerator processors.

We demonstrate that the array syntax of Fortran maps surprisingly well
onto GPUs when transformed to OpenCL kernels.  These Fortran language
features include pure and elemental functions and array constructs
like where and shift.  In addition we add a few functions that enable
a program to be deployed on machines with a hierarchy of processing
elements, such as nodes employing GPU acceleration, \emph{without
  requiring explicit declaration of parallelism within the program.}
In addition the program uses entirely standard Fortran so it can be on
a single core without concurrency.  This work also is applicable
to vendor-specific languages similar to OpenCL such as the NVidia
CUDA language.

We provide (via Fortran interfaces in the ForOpenCL library) a mechanism to
call the C OpenCL runtime and enable Fortran programmers to access OpenCL
kernels.  Transformations are supplied that provide a mechanism for converting
Fortran procedures written in the Fortran subset described in this paper to
OpenCL kernels.  We use the ROSE compiler infrastructure \cite{} to develop
these transformations.  ROSE uses the Open Fortran Parser \cite{} to parse
Fortran 2008 syntax and can output (unparse) to OpenCL.  Since ROSE's
intermediate representation (IR) was constructed to represent multiple
languages, it is relatively straightforward to transform high-level Fortran IR
nodes to OpenCL nodes.  Furthermore, the Fortran array syntax maps directly to
one, two, and three-dimensional thread groups in OpenCL.

Transformations for general Fortran procedures are not provided.  Furthermore,
a mechanism to transform the calling site to automatically invoke
OpenCL kernels is not provided at this time.  While it is possible to
accomplish this task within ROSE, it is considered outside the scope of this
paper.

We examine the performance of the Fortran data-parallel abstraction when
transformed to OpenCL to run on GPU architectures.  Since single node performance
is often given as a reason for not using data-parallel constructs within Fortran,
we consider the performance of serial data-parallel codes compared with the
usage of explicit loop constructs.

We study automatic transformations and the performance for an
application example that is typical of many applications that are based on
finite-difference or finite-volume methods in computational fluid dynamics
(CFD).  The example is a simple shallow water model in two dimensions using
finite volume methods with .... for time update.

An initial study was made for an important procedure in PAGOSA, a
non-research, production-grade code at LANL completely written in
data-parallel Fortran.  We investigated automatically transforming
this code to run on LANL's Petaflop Roadrunner computer (a hybrid
mixture of AMD Opterons and IBM Cell processors).  We demonstrated that a
source-to-source compiler can automatically vectorize and parallelize
(both within the Cell processor and across nodes) a small section of
this code.  Preliminary results showed a 12 times performance gain on
the Cell processor over a traditional single core processor.  A factor
of 9 is gained from vectorizing and parallelizing the code to run on
the Cell and the rest is gained from automatically overlapping
inter-node communication with computation on the Cell.

%%We study automatic transformations of Fortran for two applications examples.  The first is a simple shallow water model in two dimensions using finite volume methods with .... for time update.  The second is based on the primary compute kernels from PetaVision, a high-performance, neural science framework that is used in models of visual cortex.

%% To simplify things, lets save PetaVision for the local stuff as it requires convolutions

\subsection{Why Fortran?}

Fortran (originally named FORTRAN, the FORmula TRANslation
language\footnote{The capitalization convention for referring the
  language was changed in 1990 to \emph{Fortran}}) is the oldest
high-level programming language in continuous use since its
introduction, and was developed to facilitate the translation of math
formulae into machine code. Fortran was the first major language to
use a compiler to translate from a high-level program representation
to assembly language. Due to its age, it carries certain arcane
baggage that later programming languages have evolved away from. As a
result, Fortran has fallen into disfavor in certain programming
circles.  Modern versions of the language standardized in 1990, 1995, 2003, and 2008
have removed much of this legacy baggage, but these changes are not
widely known.  Modern Fortran exhibits features that are similar to 
other modern programming languages, and does not mandate legacy features
from decades old, deprecated versions of the language.

With the introduction of Fortran 90 (and later revisions to the
standard), Fortran became a truly modern programming language. Much of
the arcane baggage associated with it is actually no longer part of
the language used by modern programmers, and exists in a deprecated
form simply to support legacy codes. It is now modular and has many
object-oriented features.  It now includes a type system in which rich
first-class array data types and corresponding syntax are part of the
language, something that languages like C continue to lack.
Furthermore, Fortran should be of interest to those studying parallel
programming because of its functional and data parallel constructs and
because of the coarray notation introduced in Fortran 2008. Unlike
languages like C and C++, which are considered to be more modern,
Fortran has become a truly parallel language with features added to
recent language standards.

Yet, a likely question that one may pose is ``\emph{Why Fortran and not a more
  modern language like X?}''  The recent rise in interest in concurrency and
parallelism at the language level driven by multicore CPUs and manycore
accelerators has driven a number of new language developments, both as novel
languages and extensions on existing ones.

%% NEW - CER
So perhaps it is time to replace Fortran with yet another computer language.  The
problems with replacing Fortran with an entirely new language are two fold:
the economics of replacing the existing application base and the difficulty in
obtaining programmer acceptance.  It is estimated that replacing a major
production application at Los Alamos National Laboratory would cost between 50
and 150 million dollars.  In terms of programmer acceptance, there is always
the "chicken and egg problem": programmers won't use a new language until they
can expect good performance across a variety of platforms, and compiler
vendors can't afford to produce quality compilers until there is a reasonable
expectation of a market.

For scientific users, these new languages and language extensions
present a challenge: how do developers effectively use them while
avoiding rewriting their code and potentially growing dependent on a
transient technology that will vanish tomorrow?

%% NEW - CER
History has also shown that an investment in rewriting code does not guarantee
success either, as seen in an effort at LANL to modernize a legacy Fortran
code with the the newer C++ POOMA framework.  This massive overhaul effort led
to a code that was both slower and less flexible than the original Fortran
\cite{}.  Similar experiences have occurred in the past, notably during the
devlopment of the Sisal language.

Fortran is unique in that it has contained language features that are
well suited to modern architectures for a number of years.  This
should be unsurprising --- Fortran was a primary language used to
target systems such as the vector supercomputers and massively
parallel systems of the 1970s and 1980s.  These are the systems in
which architectural features were developed that have led to single
chip high performance architectures of interest today.  Given that
these new systems have features very similar to their predecessors, it
is clear that the language features within Fortran for them are still
relevant.

\subsection{Comparison to Other Languages}

A number of previous efforts have exploited data parallel programming
at the language level to utilize novel architectuers, particularly
in previous decades during the reign of vector and massively
parallel computers in the high performance computing world.  The
origin of the array syntax that was adopted in Fortran 90 can be found
in the APL language\cite{iverson79apl}.  Fortran 90 differed from
previous extensions of Fortran in that parallelism within whole-array operations
was expressed at the expression level instead of via parallelism within explicit
DO-loops (such as within IVTRAN for the Illiac IV).

The High Performance Fortran (HPF) extension of Fortran 90 was
proposed to add features to the language that would enhance the
ability of compilers to emit fast parallel code for distributed and
shared memory parallel computers\cite{koelbel94hpf}.  One of the
notable additions to the language in HPF was syntax to specify the
distribution of data structures amongst a set of parallel processors.
HPF also introduced an alternative looping construct to the
traditional DO-loop called \emph{FORALL} that was better suited for
parallel compilation.  An additional keyword, \emph{INDEPENDENT}, was
added to allow the programmer to indicate when the order of execution
of the program (such as a sequence of loop iterations) can be flexible
in order to allow parallel execution.  Interestingly, the parallelism
features introduced in HPF did not exploit the new array features
introduced in 1990 in any significant way, relying instead on explicit
loop-based parallelism.  This was likely in order to support parallel
programming that wasn't easily mapped onto a pure data parallel model.

In some instances though, a purely data parallel model is appropriate
for part or all of the major computations within a program.  One of
the systems where programmers relied heavily on higher level
operations instead of explicit looping constructs was the Thinking
Machines Connection Machine 5 (CM-5).  A common programming pattern
used on the CM-5 that we exploit in this paper was to write
whole-array operations from a global perspective in which computations
are expressed in terms of operations over the entire array instead of
a single local index.  The use of the array shift intrinsic functions
(like {\tt CSHIFT}) were used to build computations in which arrays
were combined by shifting the entire arrays instead of working based
on local offsets from singe indices.  A simple 1D example is one in
which an element is replaced with the average of its own value with
that of its two direct neighbors.  Ignoring boundary indices that wrap
around, explicit indexing will result in a loop such as:

\begin{verbatim}
do i = 2,(n-1)
  X(i) = (X(i-1) + X(i) + X(i+1)) / 3
end do
\end{verbatim}

When shifts are employed, this can be expressed as:

\begin{verbatim}
X = (cshift(X,-1) + X + cshift(X,1)) / 3
\end{verbatim}

Similar whole array shifting was used in higher dimensions for finite
different codes within the computational physics community, especially
at Los Alamos for codes targetting the CM-5 system that resided there until the
late 1990s.

The whole-array model was attractive because it deferred
responsibility for optimally implementing the computations to the
compiler.  Instead of relying on a compiler to infer parallelism from
a set of explicit loops, the choice for how to implement loops was
left entirely up to the tool.  Unfortunately, this had two side
effects that have limited broad acceptance of the whole-array
programming model in Fortran.  First, programmers must translate their
algorithms into a set of global operations.  Finite difference
stencils and similar computations are traditionally defined in terms
of offsets from some central index.  Shifting, while conceptually
analogous, can be awkward to think about for high dimensional stencils
with many points.  Second, the semantics of these operations are such
that all elements of an array operation are updated as if they were
updated simultaneously.  In a program where the programmer explicitly
manages arrays and loops, double buffering techniques and user managed
temporaries are used to maintain these semantics.  When the compiler
is responsible for managing this intermediate storage, it has
historically proven that they are inefficient and generate code that
requires far more temporary storage than really necessary.  This is
not a flaw of the language constructs, but a sign of the lack of
sophistication of the compilers with respect to their internal
analysis to determine how to optimally generate this intermediate
storage.

An interesting line of language research that grew out of the early work
with HPF was that associated with the ZPL language work at the University
of Washington~\ref{chamberlain04zpl}.  In ZPL, programmers adopt a simimlar
global view of computation over arrays, but define their computations based
on the local view of indices that participate in the update of each element of
an array.  For example, in the 1D averaging case stated above, a programmer would
define the computation in terms of the update that occurs at each index in terms
of relative offsets from that index:

\begin{verbatim}
Write some ZPL
\end{verbatim}

This would then be executed over the entire array. {\bf more...}

Stuff on stencil compilers?  No?