### 2010-12-14 opencl/papers/source-to-source.tex
Backup copy

\section{Source-To-Source Transformations}

This section describes the transformations that take Fortran elemental and pure procedures
as input and generate OpenCL kernels.

\subsection{OpenCL}

OpenCL~\cite{opencl:standard} is an open language standard for developing
applications for accelerators.  The C-based language provides extensions for
programming kernels that run on accelerator processing elements.  The kernels
are run by calling a C runtime library from the OpenCL host (normally the
CPU).  Efforts to standardize a C++ runtime are underway and Fortran
interfaces to the C runtime are described later.

An important concept in OpenCL is that of a thread and a thread group.  Thread
groups are used to run an OpenCL kernel concurrently on several processor
elements on the OpenCL device (often a GPU).  Consider a data-parallel
statement written in terms of an elemental fuction as discussed above.  The
act of running an OpenCL kernel can be thought of as having a particular
thread assigned to each instance of the call to the elemental function as it
is mapped across the arrays in the data-parallel statement.  In practice,
these threads are packaged into thread groups when they are run on the device
hardware.

Device memory is separated hierarchically.  A thread instance has access to
its own memory, thread groups to OpenCL local memory, and all thread groups
have access to OpenCL global memory.  When multiple members of a thread group
access the same memory elements, for example if {\tt region} or {\tt shift}
functions are called, for performance reasons it is often best if global
memory accessed by a thread group is copied into local memory.

The \emph{region} and \emph{halo} constructs easily map onto the OpenCL memory
hierarchy.  A schematic of this mapping is shown in Figure~\ref{fig:cl-memory}
for a two-dimensional array with a 2x2 array of 4 thread groups.  The memory
for the array and its halo are stored in global memory on the device as shown
in the background layer of the figure.  The array copy in local memory is
shown in the foreground divided into 4 \emph{local} tiles that partition the
array.  Halo regions in global memory are shown in dark gray and halo regions
in local memory are shown in light gray.

We emphasize that the hierarchical distribution of memory used on the
OpenCL device shown in Figure~\ref{fig:cl-memory} can be extended to include memory
across MPI nodes as well.  In this case, the virtual global array is represented by the background
layer (with its halo) and its partitions stored in the 4 MPI nodes shown in the
foreground.

Halo regions are constrained semantically so that they can not be written to
by an OpenCL kernel because the \emph{region} and \emph{interior} functions
return copies of the global memory.  Thus once memory for a halo region has
been transferred into global device memory by all of the host nodes running
MPI (before the OpenCL kernel is run), memory is in a consistent state so that
the kernels are free to read from global device memory.  Because the local
memory is a copy, it functions as a software cache for the local thread group.
Thus the compiler must insert OpenCL barriers at proper locations in the code
to insure that all threads have written to the local memory cache before any
thread can read from the cache.  On exit from a kernel, the local memory cache
is copied back to global memory for all \emph{interior} regions leaving global
memory in a consistent state again.

\subsection{ForOpenCL}

This work describes transformations that automatically create OpenCL kernel
functions from Fortran pure and elemental procedures.  These transformations
will not transform an entire program.  Users, for now, must explicitly replace
calls to Fortran kernel procedures that run on the device, with calls to the
OpenCL runtime that will run the kernel on the OpenCL host.  While these host
transformations are straightforward using ROSE, they are outside the scope of
this paper.

In addition to the Fortran to OpenCL transformations, the
ForOpenCL~\cite{http:ofp} library provides programmers with the ability to
call the C OpenCL runtime from Fortran.  ForOpenCL is a set of Fortran modules
providing Fortran 2003 interface descriptions and classes that allow language
interoperability with the OpenCL runtime.

\section{Transformation examples}

This sections show short Fortran code examples and OpenCL equivalent.  The
notation uses uppercase for arrays and lowercase for scalar quantities.
Interior and region array copies are denoted by an i or an r preceeding the array.
For example, {\tt iH = interior(H, halo)} is a local copy of interior region
of array {\tt H} (representing height) in the shallow water code.

\subsubsection{interior and region functions}

While the serial versions of the interior and regions functions return an
array copy, in OpenCL, these functions return a scalar quantity based on
the location of a thread in a thread group and the relationship of its location
to the array copy in local memory.  Because we assume there is a thread for
every element in the interior, the array index is just the thread index adjusted
for the halo.  Thus {\tt interior} and {\tt region} are just inline OpenCL
functions provided by the ForOpenCL library.

\subsubsection{array syntax}

\begin{verbatim}
   V = V - iA*(V - v_rest)                  !! Fortran

   V[l] = V[l] - iA[lex]*(V[l] - v_rest);   // OpenCL equivalent
\end{verbatim}

\subsubsection{where construct}

A neuron in PetaVision fires (activity is set to 1)
whenever the action potential is greater than some threshold.
This is easily expressed with a where construct.

\begin{verbatim}
   where (V > Vth)                        !! Fortran
      iA = 1
   elsewhere
      iA = 0
   end where

   iA[lex] = (V[l] > Vth[l]) ? 1 : 0;     // OpenCL equivalent
\end{verbatim}

\subsection{New functions}

%\subsubsection{transfer_halo}
\subsubsection{region}

One of the state variables in the shallow water code is H, effectively the
height of the water.  This variable has a halo (ghost-cell region) surrounding
the interior of the grid to handle boundary conditions.  When using
MPI, the halo regions contains values from adjacent processors that must be
updated with new values at each time step.  This is accomplished with the
{\tt transfer\_halo} function.  The shallow water code assumes a five point stencil
so the state variables are extended by 2 in each dimension (e.g., by one to the
left, right, up, and down).

\begin{verbatim}
halo = [1,1,1,1]
iH => region(H, halo, transfer=false)
\end{verbatim}