### 2010-12-14 opencl/papers/shallow-water.tex
Initial version

\section{Shallow Water Model}

A short description of the shallow water model goes here.

\subsection{Serial code examples}

The Fortran kernel procedure {\tt wave_advance} for the shallow water code
is declared as:
\begin{verbatim}
   pure subroutine wave_advance(H, U, V, dx, dy, dt)
      real, dimension(:,:), intent(inout) :: H, U, V
      real, intent(in) :: dx, dy, dt
      !$OFP CONTIGUOUS :: H, U, V
      !$OFP KERNEL :: wave_advance
   end subroutine
\end{verbatim}
where {\tt H, U}, and {\tt V} are start variables for height and
$x$ and $y$ momentum respectively.  The \emph{OFP} compiler directives
{\tt CONTIGUOUS} and {\tt KERNEL} are added because contiguous is Fortran
2008 syntax (not in current compilers) and kernel is an \emph{OFP} extension.

Temporary arrays are required for the interior copies {\tt iH, iU}, and
{\tt iV} of the state variables and for the flux quantities {\tt Hx, Hy, Ux,
Vx}, and {\tt Vy} defined on cell faces.  These temporary arrays are declared as,
\begin{verbatim}
   real, allocatable, dimension(:,:) :: iH, iU, iV
   real, allocatable, dimension(:,:) :: Hx, Hy, Ux, Uy, Vx, Vy
\end{verbatim}

Halos variables for the interior and cell faces are defined as
\begin{verbatim}
   integer, dimension(4) :: halo, face_lt, face_rt, face_up, face_dn
   halo    = [1,1,1,1]
   face_lt = [0,1,1,1]; face_rt = [1,0,1,1]
   face_dn = [1,1,0,1]; face_up = [1,1,1,0]
\end{verbatim}
Note that the halos for the four faces each have a 0 in the definition. Thus the returned
array copy will have size that is larger than the interior regions that use {\tt [1,1,1,1]}.
This is because there is one more cell face quantity than there are cells in a given direction.

The Lax Wendrof (???) update algorithm for the shallow-water model has two steps.
The fluxes are first updated on the cell faces, for example,
\begin{verbatim}
   Hx =           0.5 * ( region(H, face_lt) + region(H, face_rt) ) &
        + (0.5*dt/dx) * ( region(U, face_lt) - region(U, face_rt) )
\end{verbatim}
updates the $x$-direction flux for the height state variable.

Then second step uses these fluxes to update the interior regions for the state varaibles.
For example,
\begin{verbatim}
   face_lt = [0,1,0,0];  face_rt = [1,0,0,0]
   face_dn = [0,0,0,1];  face_up = [0,0,1,0]
   iH = iH + (dt/dx) * ( region(Ux, face_lt) - region(Ux, face_rt) ) &
           + (dt/dy) * ( region(Vy, face_dn) - region(Vy, face_up) )
\end{verbatim}
Note that face halos have been redefined so that the array copy returned by
now has the same size as the interior region.

These simple code segments show how the shallow water model is
implemented in standard Fortran (2003) using the data-parallel
programming model described above.  The resulting code is simple,
consise, and easy to understand.  However it does \emph{not}
necessarily perform well because of the temporary array variables,
especially those produced by the {\tt region} function.  This is
generally true of algorithms that use Fortran shift functions as well,
as some Fortran compilers (e.g., gfortran) do not generate optimal
code for shifts.  We note (as seen below) that these temporary array
copies are replaced by scalars in the transformed Fortran code so
there is no performance penalty for using data-parallel statements
as outlined.