### 2010-12-15 opencl/papers/programming-model.tex
edits

\section{Programming Model}

The static analysis and source-to-source transformations used are very
basic and only require the programmer to use a subset of Fortran that
employs a data-parallel programming model.  In particular, it
encourages use of language features that were introduced and
standardized in the Fortran 90 language specification.  In this
section we describe the set of four Fortran 90 features that our analysis
and transformation method are based on.  From these language constructs,
we are able to easily transform them to a lower-level CUDA or OpenCL
implementation.

%This paper uses four simple features of Fortran that form a language subset
%that does not require any language extensions and can be easily transformed
%to a lower-level implementation in either CUDA or OpenCL.

\subsubsection*{Array notation}

Fortran 90 introduced a rich array syntax that allows programmers to
write statements that are in terms of whole arrays or subarrays, with
data parallel operators to compute on the arrays.  This has the
benefit of avoiding explicit looping in the code and maintaining a
high-level style that is closer to the original mathematical
specification of the problem being solved.  More importantly from a
compilation perspective, this defers decisions about how to implement
these whole-array operations to a compilation tool.  Sophisticated
loop analysis to identify parallelism within the loop are not
necessary.  When faced with a novel architecture such as modern GPUs
that are ideally suited to data parallel programming models, the fit
between Fortran arrays and these systems is quite clean.

%% NEW - CER
Arrays are first class citizens in Fortran instead of a basic pointer
to raw memory as in many other compiled languages.  A Fortran array
contains metadata associated with it that describes the array's shape
and size. It should be noted that other languages that provide rich
array data types and whole array operations may also be suitable
targets for the transformations described in this paper.  
In addition to metadata about the structure of the arrays, operators
are provided that allow programmers to write expressions based on the
whole array instead of explicit indexing and application of operations
to individual elements.  For example, if {\tt A}, {\tt B}, and {\tt C}
are all arrays of the same rank and shape and {\tt s} is a scalar,
then the statement

{\small
\begin{verbatim}
 C = A + s*B
\end{verbatim}
}

results in the element-wise sum of {\tt A} and {\tt s} times the elements of
{\tt B} being stored in the corresponding elements of {\tt C}. The first
element of {\tt C} will contain the value of the first element of {\tt A}
added to the first element of {\tt c*B}.  Note that no explicit iteration over
array indices is needed and that the individual operators, plus, times, and
assignment are applied by the compiler to individual elements of the arrays
independently.  Thus the compiler is able to spread the computation in the
example across any hardware threads under its control.

%% NEW - CER
%%% The use of array notation allows one to program in a data-parallel subset of Fortran. This style of programming makes use of array notation and pure and elemental functions to operate on array elements.

\subsubsection*{Elemental functions}

An elemental function consumes and produces scalar values, but can be applied
to variables of array type such that the function is applied to each and every
element of the array.  This allows programmers to avoid explicit looping and
instead simply state that they intend a function to be applied to every
element of an array in parallel, deferring the choice of implementation
technique to the compiler.  Elemental functions are intended to be used for
data parallel programming, and as a result must be side effect free (or,
\emph{pure}).

%% NEW - CER
%%% Because elemental functions return scalar values and are free from side effects, the compiler is free to distribute the computation over any hardware processing elements available to it, such as the multiple cores and vector units on an Intel or AMD processor or the Synergistic Processing Units on an IBM Cell processor.

%% NEW - CER
For example, the basic array operation shown above could be refactored into
an elemental function,

{\small
\begin{verbatim}
  pure elemental real function foo(a, b, s)
    real, intent(in) :: a, b, s
    foo = a + s*b
  end function
\end{verbatim}
}

\noindent and called with

{\small
\begin{verbatim}
  C = foo(A, B, s)
\end{verbatim}
}

Note that while {\tt foo} is defined in terms of purely scalar
quantities, it can be \emph{applied} to arrays as shown.  While this
may seem like a trivial example, such simple functions may be composed
with other elemental functions to perform powerful computations,
especially when applied to arrays.  Our prototype tool transforms
elemental functions to inline OpenCL functions.  Thus there is no
penalty for usage of elemental functions and provide a convient
mechanism to express algorithms in simpler segments.

%Also note that the plus and times operators shown in the data-statement
%example can be seen as an elemental functions, as plus and times are defined
%in terms of scalars but can be applied to whole arrays and can return an array
%result.

\subsubsection*{Pure procedures}

Pure procedures, like elemental functions, must be free of side
effects.  Unlike elemental functions that require arguments to have an
{\tt intent(in)} attribute, they may change the contents of array
arguments that are passed to them.  The absence of side effects
removes ordering constraints that could restrict the freedom of the
compiler to invoke pure functions out of order and possibly in parallel.
Procedures and functions of this sort are also common in pure
functional languages like Haskell, and are exploited by compilers in
order to emit parallel code automatically due to their suitability for
compiler-level analysis.

Since pure procedures don't have side effects they a candidates for
running on accelerators in OpenCL.  Currently our prototype tool
transforms pure procedures to OpenCL kernels that \emph{do not} call
other procedures, except for elemental functions.

\subsubsection*{Shift functions}

Many array-based algorithms require the same operation to be performed
on each element of the array using the value of that element and some small
set of neighboring cells.  Often programmers implement these operations that
are local to each element within the inner-loop of a set of nested FOR-
or DO-loops using offsets relative to the current array index.  An alternative
to this local-view of the algorithm is to take a global view and write the
algorithm in terms of the whole array.  For example, consider a 1D array in
which we wish to subtract the $(i-1)$th element from the $i$th for all
elements.  One way to look at this is that we are subtracting the entire array
shifted by one element from itself.

Fortran provides a set of shifting operators that allow programmers to
define operations based on shifted arrays.  These intrinsic operators take
an array, a dimension, and the amount by which it should be shifted (using
the sign to indicate direction).  By defining operations on entire arrays
based on a global view of them shifted relative to each other, programmers can
avoid explicit looping and potentially tricky index arithmetic.  Furthermore,
analysis of the extent of the set of shifted arrays in a given expression
allows analysis tools to determine the amount of temporary or buffer storage
necessary to hold intermediate values during whole array operations.  With
explicit loops, programmers must maintain this temporary storage manually.

\subsubsection*{Regions}

Borrowing from ZPL, we introduce the concept of regions to Fortran.  In
Fortran, regions are expressed as pointers to a subsection of an existing array.
Regions may refer to an interior portion of an array or the entire array.

The use of regions implies the use of the halo (or ghost) cell pattern 
where the size of an array is increased to provide extra array elements
surrounding the interior portion of the array.  
%parlab.eecs.berkeley.edu/wiki/_media/patterns/ghostcell.pdf

Regions are similar to the shift operator as they can be used to reference
portions of the array that are shifted with respect to the interior portion.
However, unlike the shift operator, regions are not expressed in terms of
boundary conditions and thus don't explicitly \emph{require} a knowledge of
nor the application of boundary conditions locally.  Thus, as will be shown
below, regions are more suitable for usage by OpenCL thread groups which
access only local subsections of an array stored in global memory.

\subsection{New Procedures}

Unlike OpenCL, we require limited use of compiler directives, although
two are used to enforce the semantics required by the programming
model.  These directives are:

\begin{itemize}

\item {\tt \!\$OFP CONTIGUOUS}: specifies that an dummy array variable is
  contiguous in memory. CONTIGUOUS is an attribute in Fortran 2008.

\item {\tt \!\$OFP KERNEL}: specifies that a pure subroutine can be
  transformed to an OpenCL kernel.

\end{itemize}

Three new procedures (in addition to the intrinsic shift function) are
defined in Fortran that are used in array-syntax operations.  Each
procedure takes a integer array halo argument that specifies the
number of ghost cells on either side of a region, for each dimension.
For example {\tt halo = [left, right, down, up]} specifies a halo for
a two-dimensional region.  These functions are:

\begin{itemize}

\item {\tt transfer\_halo(array, halo)}: an impure subroutine that
  exchanges halo regions between nodes using MPI or Fortran
  coarrays. Array is the memory stored on the node representing the
  node-local portion of a virtual global array. Not used in this work.

\item {\tt interior(array, halo)}: a pure function that returns a copy of
  the interior portion of the array specified by halo. Array is
  local to a node.

\item {\tt region(array, halo)}: a pure function that returns a copy of
  the portion of the array specified by halo.  Array is local to a node.

\end{itemize}

It should be noted that the two functions {\tt interior} and {\tt
  region} are pure and thus can be called from within a pure kernel
procedure.  These two functions are part of the language recognized by
the compiler and though the two functions return a copy of a portion
of an array \emph{semantically}, the compiler is not forced to
actually make a copy and is free to enforce copy semantics through
other means.

\subsection{Parallelism}

% TODO - Matt - do you want a try at this.  Not sure we have room for
% this subsection. I briefly mention node parallelism in other sections.

% need to describe processor hierarchy of nodes and accelerators
% need to describe memory hierarchy as well in regards to global array view,
% node array view (with a copy is GPU global memory) and local GPU memory which
% further subdivides the node array memory

% figures on slicing and dicing of memory region go here

There are several advantages to this style of programming using array
syntax, shifts, regions, and pure and elemental functions:

\begin{itemize}
\item There are no loops or index variables to keep track of.  Off by
  one index errors and improper handling of array boundaries are a
  common programming mistake.
\item The written code is closer to the algorithm, easier to
  understand, and is usually substantially shorter.
\item Semantically the intrinsic functions return arrays by value.
  This is usually what the algorithm requires.
\item Because pure and elemental function are free from side effects,
  it is easier for a compiler to schedule the work to be done in
  parallel.
\end{itemize}

%% MJS - mention that these features were designed for parallelism in
%% the first place?

%% NEW - CER
An example of this style of programming in Fortran is shown in

{\small
\begin{verbatim}
  Bz = Bz &
     + dt*(cshift(Ex,dim=2,shift=+1)-Ex)/dy &
     - dt*(cshift(Ey,dim=1,shift=+1)-Ey)/dx
\end{verbatim}
}

%% NEW - CER

This example is a solution to Maxwell's equations for the $z$
component of the magnetic field using Fortran array syntax.  Note that
there are no explicit loops in this example.  The operators {\tt +},
{\tt -}, and {\tt *} are applied to all of the elements of the
three-dimensional arrays {\tt Bz}, {\tt Ex}, and {\tt Ey},
individually.  This is why data parallelism has been called
collection-oriented programming by
Blelloch~\cite{blelloch90,rajopadhyedidlacs}.  As the {\tt cshift}
function and the array-valued expressions all semantically returns a
value, this style of programming is also similar to functional
programming (or value-oriented
programming).

%%~\cite{simonpeytonjones}).  The heart of the solution
%to Maxwell's equations is the statement shown in
%Listing~\ref{lst:dpexample} and five similar, simple equations.

%% NEW - CER
%%% The power in this notation is that the compiler is free to distribute the computation in these expressions over any hardware processing elements available to it, such as the vector units on an Intel or AMD processor or the Synergistic Processing Units on an IBM Cell processor.  If the arrays are declared as coarrays, this includes spreading the computation over \emph{all of the nodes in   a cluster as well.}  In this case, communication occurs within the \texttt{cshift} functions, though the compiler is free to overlap communication and computation by scheduling the communication early and performing the computation on the interior of the arrays while waiting for the communication to complete.

%% NEW - CER
Complete and very concise and elegant programs can be built with procedures
similar to the example shown above. To aid this effort, Fortran
supplies intrinsic functions like the array constructors (\texttt{CSHIFT},
\texttt{EOSHIFT}, \texttt{MERGE}, \texttt{TRANSPOSE}, ...), the array location
functions (\texttt{MAXLOC} and \texttt{MINLOC}), and the array reduction
functions (\texttt{ANY}, \texttt{COUNT}, \texttt{MINVAL}, \texttt{SUM},
\texttt{PRODUCT}, ...).  To this set we add region functions described above.

%% NEW - CER
This style of programming meets the requirements we have set for a programming
model for developing applications suitable for acceleration.  It allows the programmer to
program at a very-high level of abstraction while providing the compiler with
maximum flexibility in targeting the application for a particular hardware
architecture.  The data parallel programming model simultaneously meets the
seemingly conflicting goals of maintainability, portability, and performance.

%% NEW - CER
Unfortunately, this style of programming has never really caught on because
when Fortran 90 was first introduced, performance was relatively poor and thus
programmers shied away from using array syntax (even now, some are actively
counseling against its usage because of performance issues~\cite{Cray-SC07}).
Thus the Fortran community was caught in a classic ``chicken-and-egg''
conundrum: (1) programmers didn't use it because it was slow; and (2)
compilers vendors didn't improve it because programmers didn't use it.

%% NEW - CER
A goal of this paper is to demonstrate that parallel programs written in
this style of Fortran are maintainable and can achieve good performance on a
variety of processor architectures, including possibly, future terascale
architectures from Intel (such as the XXX architecture).

\subsection{Limitations}

Only Fortran procedures are transformed into OpenCL kernels.  The programmer
must currently explicitly call these kernels from Fortran using the ForOpenCL
library described below.  It is also possible using ROSE to modify the calling
site so that the entire program can be transformed but this functionality is
outside the scope of this paper.  Here we specifically examine transforming
Fortran procedures to OpenCL kernels.

Only elemental functions may be called from kernel functions.  These include
fortran functions that have an OpenCL analog and user-defined elemental functions.

A kernel procedure (specified by the {\tt \!\$OFP KERNEL} directive) must be
pure and all array variables must be declared as contiguous.  A kernel
procedure may not call other procedures except for limited intrisic
functions (primarily math), user-defined elemental function, and the
{\tt interior} and {\tt region} functions.

Array sizes must be multiples of the local kernel size, {\tt
  get\_local\_size(0)*get\_local\_size(1)}.  This may be relaxed in
the future.