### 2010-12-20 opencl/papers/foropencl.tex
Nearly final draft.

%
% I'm adding an outline here so I can see how everything fits together

%
% \title{Exploiting First-Class Arrays in Fortran for Accelerator Programming}
%

%
% \abstract
%

%
% \section{Introduction}
%    \subsection{Approach}
%    \subsection{Why Fortran?}
%    \subsection{Comparison to Other Languages}
%

%
% \section{Programming model}
%    \subsection{Fortran Syntax}   
%       \subsubsection*{Array notation}
%       \subsubsection*{Elemental functions}
%       \subsubsection*{Pure procedures}
%       \subsubsection*{Shift functions}
%       \subsubsection*{Regions}
%    \subsection{New Functions}
%    \subsection{Parallelism}
%    \subsection{Limitations}
%

%
% \section{Shallow Water Model}
%    \subsection{Equations}
%

%
% \section{Source-To-Source Transformations}
%    \subsection{ForOpenCL}
%       \subsubsection{array syntax}
%       \subsubsection{where construct}
%    \subsection{New functions}
%       \subsubsection{region}
%    \subsection{Static Analysis}
%       \subsubsection{Analysis not required}
%    \subsection{Simplifying Assumptions}
%

%
% \section{Performance}
%

%
% \section{Conclusions}
%

\documentclass[10pt, conference, compsocconf]{IEEEtran}

\usepackage{cite}
\usepackage{graphicx}
\usepackage[cmex10]{amsmath}
\usepackage{amssymb}
\usepackage{array}
\usepackage{url}
\usepackage{listings}

\lstdefinestyle{FortranLike}{float,frame=lines,language=Fortran,commentstyle=\ttfamily,basicstyle=\ttfamily}
\lstdefinestyle{CLike}{float,frame=lines,language=C,commentstyle=\ttfamily,basicstyle=\ttfamily}
\lstdefinestyle{NoFloatCLike}{frame=lines,language=C,commentstyle=\ttfamily,basicstyle=\ttfamily}

\title{Exploiting Array Syntax in Fortran for Accelerator Programming}

%%
%% reorder appropriately later
%%
\author{\IEEEauthorblockN{Matthew J. Sottile}
\IEEEauthorblockA{Galois, Inc.\\
%421 SW 6th Ave. Suite 300 \\
%Portland, OR 97204\\
Email: matt@galois.com}
\and
\IEEEauthorblockN{Craig E Rasmussen,\\
                  Wayne N. Weseloh,\\
                  Robert W. Robey}
\IEEEauthorblockA{Los Alamos National Laboratory\\
%CCS-7, MS B287\\
%Los Alamos, NM\\
Email: crasmussen@lanl.gov}

\and
\IEEEauthorblockN{Daniel Quinlan}
\IEEEauthorblockA{Lawrence Livermore\\ National Laboratory}

\and
\IEEEauthorblockN{Jeffrey Overbey}
\IEEEauthorblockA{Indiana University}

}

\begin{document}

\maketitle

\begin{abstract}
Emerging architectures for high performance computing are well suited
to a data-parallel programming model.  This paper presents a simple
programming methodology that allows programmers to take advantage of
these emerging systems.  We use array constructs in Fortran 90 to show how this
infrequently exploited, standardized language feature is easily
transformed to lower-level accelerator code.  These transformations
are based on a simple mapping from Fortran 90 to OpenCL.  We
demonstrate, using a shallow-water code, that an acceleration of up to 40
times can be achieved on GPU hardware with these compiler-generated
transformations.
\end{abstract}

\input{intro}
\input{programming-model}
\input{shallow-water}
\input{source-to-source}
%\input{static-analysis}
\input{performance}

\section{Conclusions}

The sheer complexity of programming for clusters of many or multi-core
processors with tens of millions threads of execution make the simplicity of
the data-parallel model attractive.  Furthermore, the increasing complexity of
todays applications (especially when convolved with the increasing complexity
of the hardware) and the need for portability across hardware architectures
make a higher-level and simpler programming model like data-parallel
attractive.

The goal of this work has been to exploit source-to-source transformations that
allow programmers to develop and maintain programs at a high-level of
abstraction, without coding to a specific hardware architecture.
Furthermore these transformations allow multiple hardware architectures
to be targeted without changing the high-level source.  It also removes the
necessity for application programmers to understand details of the accelerator
architecture or to know OpenCL.

%%TODO \cite{chamberlain04zpl, roth97stencils}

\bibliographystyle{IEEEtran}
\bibliography{IEEEabrv,foropencl}

\end{document}