### 2010-12-20 opencl/papers/source-to-source.tex
tweaks to shrink to 8 pages

\section{Source-To-Source Transformations}

This section describes the transformations that take Fortran elemental and
pure procedures as input and generate OpenCL kernels.

\subsection{OpenCL}

OpenCL~\cite{opencl08} is an open language standard for developing
applications for accelerators.
The kernels
are run by calling a C runtime library from the OpenCL host (normally the
CPU).  Efforts to standardize a C++ runtime are underway and Fortran
interfaces to the C runtime are described later.

An important concept in OpenCL is that of a thread and a thread group.  Thread
groups are used to run an OpenCL kernel concurrently on several processor
elements on the OpenCL device (often a GPU).  Consider a data-parallel
statement written in terms of an elemental function as discussed above.  The
act of running an OpenCL kernel can be thought of as having a particular
thread assigned to each instance of the call to the elemental function as it
is mapped across the arrays in the data-parallel statement.  In practice,
these threads are packaged into thread groups when they are run on the device
hardware.

Device memory is separated hierarchically.  A thread instance has access to
its own memory, thread groups to OpenCL local memory, and all thread groups
have access to OpenCL global memory.  When multiple members of a thread group
access the same memory elements (if {\tt region} or {\tt shift}
functions are called), for performance reasons it is often best if global
memory accessed by a thread group is copied into local memory.

The \emph{region} and \emph{halo} constructs easily map onto the OpenCL memory
hierarchy.  A schematic of this mapping is shown in Figure~\ref{fig:cl-memory}
for a two-dimensional array with a 2x2 array of 4 thread groups.  The memory
for the array and its halo are stored in global memory on the device as shown
in the background layer of the figure.  The array copy in local memory is
shown in the foreground divided into 4 \emph{local} tiles that partition the
array.  Halo regions in global memory are shown in dark gray and halo regions
in local memory are shown in light gray.

\begin{figure}[!t]
\centering
\includegraphics[width=1.75in]{cl-memory.pdf}
\caption{A schematic of global memory for an array and its copy stored in local memory
for four thread groups.}
\label{fig:cl-memory}
\end{figure}

We emphasize that the hierarchical distribution of memory used on the OpenCL
device shown in Figure~\ref{fig:cl-memory} can be extended to include memory
across MPI nodes as well.  In this case, the virtual global array is
represented by the background layer (with its halo) and its partitions stored
in the 4 MPI nodes shown in the foreground.

Halo regions are constrained semantically so that they can not be written to
by an OpenCL kernel because the \emph{region} and \emph{interior} functions
return copies of the global memory.  Thus once memory for a halo region has
been transferred into global device memory by all of the host nodes running
MPI (before the OpenCL kernel is run), memory is in a consistent state so that
the kernels are free to read from global device memory.  Because the local
memory is a copy, it functions as a software cache for the local thread group.
Thus the compiler must insert OpenCL barriers at proper locations in the code
to insure that all threads have written to the local memory cache before any
thread can read from the cache.  On exit from a kernel, the local memory cache
is copied back to global memory for all \emph{interior} regions leaving global
memory in a consistent state again.

%\subsection{ForOpenCL}

%This work describes transformations that automatically create OpenCL kernel
%functions from Fortran pure and elemental procedures.  At present, these transformations
%will not transform an entire program.  Users, for now, must explicitly replace
%calls to Fortran kernel procedures that run on the device, with calls to the
%OpenCL runtime that will run the kernel on the OpenCL host.  While these host
%transformations are straightforward using ROSE, they are outside the scope of
%this paper.

%In addition to the Fortran to OpenCL transformations, the
%ForOpenCL library available from the Open Fortran Parser web page provides programmers with the ability to
%call the C OpenCL runtime from Fortran.  ForOpenCL is a set of Fortran modules
%providing Fortran 2003 interface descriptions and classes that allow language
%interoperability with the OpenCL runtime.

\section{Transformation examples}

This section outlines the OpenCL equivalent syntax for portions of
the Fortran shallow-water code described in
Section~\ref{sec:shallow-water}.  The notation uses uppercase for
arrays and lowercase for scalar quantities.  Interior variables are
denoted by an {\tt i} preceding the array name.  For example, {\tt iH =
  interior(H, halo)} is a local copy of interior region of array {\tt
  H} (representing height) in the shallow water model.

\subsubsection{Interior and region functions}

While the serial versions of the interior and regions functions return an
array copy, in OpenCL, these functions return a scalar quantity based on the
location of a thread in a thread group and the relationship of its location to
the array copy in local memory.  Because we assume there is a thread for every
element in the interior, the array index is just the thread index adjusted for
the halo.  Thus {\tt interior} and {\tt region} are just inline OpenCL
functions provided by the ForOpenCL library.

\subsubsection{Function and variable declarations}

Fortran kernel procedures have direct correspondence with OpenCL equivalents.
For example, {\tt wave\_advance} is transformed as {\tt \_\_kernel void
  wave\_advance(\_\_ global float * H, ..., float dt);}.  The global state
variables have local equivalents, e.g., {\tt \_\_local float
  H\_local[MAX\_LOCAL\_SIZE]}, declared with the appropriate size and are
copied to local memory by the compiler with inlined library functions.  The
array temporaries defined on cell faces are declared similarly while interior
variables are simple scalars, e.g., {\tt float iH;}.  Region variables cannot
be scalar objects because regions may be shifted and thus \emph{shared} by
threads within a thread group.

\subsubsection{Array syntax}

Array syntax transforms nearly directly to OpenCL code.  For example, interior
variables are particularly straightforward as they are scalar quantities in
OpenCL,

{\small
\begin{verbatim}
iH = iH + (dt/dx) * (region(Ux, face_lt) - 
                     region(Ux, face_rt) )
        + (dt/dy) * (region(Vy, face_dn) - 
                     region(Vy, face_up) );
\end{verbatim}
}

\noindent Region variables are more complicated because they are
arrays.

{\small
\begin{verbatim}
Hx[i] = 0.5 * (region(H_local, face_lt)+ ...);
\end{verbatim}
}

\noindent
where {\tt i = LX + LY*(NLX+halo(0)+halo(1)))} is a local index variable, {\tt
  LX = get\_local\_id(0)} is the local thread id in the $x$ dimension, {\tt LY =
  get\_local\_id(1)} is the local thread id in the $y$ dimension, {\tt NLX =
  get\_local\_size(0)} is the size of the thread group in the $x$ dimension, and
the {\tt get\_local\_id} and {\tt get\_local\_size} functions are defined by the
OpenCL language standard.