### 2011-06-23 papers/GPUScA-2011/programming-model.tex
Changes on Thursday.

\section{Programming Model}

A question that one may pose is ``\emph{Why Fortran and not a more modern
  language like X?}''  The recent rise in interest in concurrency and
parallelism at the language level driven by multicore CPUs and manycore
accelerators has driven a number of new language developments, both as novel
languages and extensions on existing ones.  For many scientific users familiar
with Fortran, new languages and language extensions to use novel new
architectures present a challenge: how do they effectively use them while
avoiding rewriting code and potentially growing dependent on a transient
technology that will vanish tomorrow?

In this section we present the Fortran subset employed in this paper.  This
subsetting language will allow scientific programmers to stay within the Fortran
language and yet have direct access to GPU hardware.  We first examine how this
programming model in Fortran relates to developments in other languages.

\subsection{Comparison to Other Languages}

A number of previous efforts have exploited data-parallel programming at the
language level to utilize novel architectures, particularly in previous decades
during the reign of vector and massively parallel computers in high performance
computing (HPC).  The origin of the array syntax adopted by Fortran in the 1990
standard can be found in the APL language~\cite{iverson79apl}.  These additions
to Fortran allowed parallelism to be expressed with whole-array operations at
the expression level, instead of via parallelism within explicit DO-loops (such
as within IVTRAN for the Illiac IV).

The High Performance Fortran (HPF) extension of Fortran was proposed to add
features to the language that would enhance the ability of compilers to emit
fast parallel code for distributed and shared memory parallel
computers\cite{koelbel94hpf}.  One of the notable additions to the language in
HPF was syntax to specify the distribution of data structures amongst a set of
parallel processors.  HPF also introduced an alternative looping construct to
the traditional DO-loop called {\tt FORALL} that was better suited for parallel
compilation.  An additional keyword, {\tt INDEPENDENT}, was added to allow the
programmer to indicate when the loop contained no loop-order dependencies that
allowed for parallel execution.  These constructs are similar to {\tt DO
  CONCURRENT}, an addition to Fortran in 2008.  Interestingly, the parallelism
features introduced in HPF did not exploit the new array features introduced in
1990 in any significant way, relying instead on explicit loop-based parallelism.
This restriction allowed the language to support parallel programming that
wasn't easily mapped onto a pure data-parallel model.

In some instances though, a purely data-parallel model is appropriate for part
or all of the major computations within a program.  One of the systems where
programmers relied heavily on higher level operations instead of explicit
looping constructs was the Thinking Machines Connection Machine 5 (CM-5).  A
common programming pattern used on the CM-5 (that we exploit in this paper) was to
write whole-array operations from a global perspective in which computations are
expressed in terms of operations over the entire array instead of a single local
index.  The use of the array shift intrinsic functions (like {\tt CSHIFT}) were
used to build computations in which arrays were combined by shifting the entire
arrays instead of working on local offsets based on single indices.  A simple
1D example is one in which an element is replaced with the average of its own
value and that of its two direct neighbors.  Ignoring boundary indices that
wrap around, explicit indexing will result in a loop such as:

{\small
\begin{verbatim}
  do i = 2,(n-1)
    X(i) = (X(i-1) + X(i) + X(i+1)) / 3
  end do
\end{verbatim}
}

\noindent When shifts are employed, this can be expressed as:

{\small
\begin{verbatim}
  X = (cshift(X,-1) + X + cshift(X,1)) / 3
\end{verbatim}
}

Similar whole array shifting was used in higher dimensions for finite
difference codes within the computational physics community, especially
at Los Alamos for codes targeting the CM-5 system that resided there until the
late 1990s.  A body of research in compilation of stencil-based codes
that use shift operators targeting these systems is related to the
work we present here~\cite{stencil-compiler}.

The whole-array model was attractive because it deferred responsibility for
optimally implementing the computations to the compiler.  Instead of relying on
a compiler to infer parallelism from a set of explicit loops, the choice for how
to implement loops was left entirely up to the tool.  Unfortunately, this had
two side effects that have limited broad acceptance of the whole-array
programming model in Fortran.  First, programmers must translate their
algorithms into a set of global operations.  Finite difference stencils and
similar computations are traditionally defined in terms of offsets from some
central index.  Shifting, while conceptually analogous, can be awkward to think
about for high dimensional stencils with many points.  Second, the semantics of
these operations are such that all elements of an array operation are updated as
if they were updated simultaneously.  In a program where the programmer
explicitly manages arrays and loops, double buffering techniques and user
managed temporaries are used to maintain these semantics.  When the compiler is
responsible for managing this intermediate storage, it has historically proven
that they are inefficient and generate code that requires far more temporary
storage than really necessary.  This is not a flaw of the language constructs,
but a sign of the lack of sophistication of the compilers with respect to their
internal analysis to determine how to optimally generate this intermediate
storage.

An interesting line of language research that grew out of the early work with
HPF was that associated with the ZPL language work at the University of
Washington~\cite{chamberlain04zpl}.  In ZPL, programmers adopt a similar global
view of computation over arrays, but define their computations based on regions,
which provide a local view of the set of indices that participate in the update
of each element of an array.  The Chapel parallel programming language being
developed by Cray is directly influenced by HPF and
ZPL~\cite{chamberlain_chapel}.

\subsection{Fortran Language Subset}

The static analysis and source-to-source transformations used in this work are
very basic and only require the programmer to use a language subset that employs
a data-parallel programming model.  In particular, it encourages the use of
array notation, elemental functions, and pure procedures.  From these language
constructs, we are able to easily transform Fortran procedures to a lower-level
OpenCL kernel implementation.

\subsubsection*{Array notation}

Fortran has a rich array syntax that allows programmers to write statements in
terms of whole arrays or subarrays, with data parallel operators to compute on
the arrays.  Array variables can be used in expressions based on whole-array
operations.  For example, if {\tt A}, {\tt B}, and {\tt C} are all arrays of the
same rank and shape and {\tt s} is a scalar, then the statement

{\small
\begin{verbatim}
 C = A + s*B
\end{verbatim}
}

\noindent results in the element-wise sum of {\tt A} and {\tt s} times the
elements of {\tt B} being stored in the corresponding elements of {\tt C}. The
first element of {\tt C} will contain the value of the first element of {\tt A}
added to the first element of {\tt c*B}.  Note that no explicit iteration over
array indices is needed and that the individual operators, plus, times, and
assignment are applied by the compiler to individual elements of the arrays
independently.  Thus the compiler is able to spread the computation in the
example across any hardware threads under its control.

\subsubsection*{Elemental functions}

An elemental function consumes and produces scalar values, but can be applied to
variables of array type such that the function is applied to each and every
element of the array.  This allows programmers to avoid explicit looping and
instead simply state that they intend a function to be applied to every element
of an array in parallel, deferring the choice of implementation technique to the
compiler.  Elemental functions are intended to be used for data parallel
programming, and as a result must be side effect free and mandate an {\tt
intent(in)} attribute for all arguments.

For example, the basic array operation shown above could be refactored into an
elemental function,

{\small
\begin{verbatim}
  pure elemental real function foo(a, b, s)
    real, intent(in) :: a, b, s
    foo = a + s*b
  end function
\end{verbatim}
}

\noindent and called with

{\small
\begin{verbatim}
  C = foo(A, B, s)
\end{verbatim}
}

Note that while {\tt foo} is defined in terms of purely scalar quantities, it
can be \emph{applied} to arrays as shown.  While this may seem like a trivial
example, such simple functions may be composed with other elemental functions to
perform powerful computations, especially when applied to arrays.  Our prototype
tool transforms elemental functions to inline OpenCL functions.  Thus there is
no penalty for usage of elemental functions and they provide a convenient
mechanism to express algorithms in simpler segments.

\subsubsection*{Pure procedures}

Pure procedures, like elemental functions, must be free of side effects.  Unlike
elemental functions that require arguments to have an {\tt intent(in)}
attribute, they may change the contents of array arguments that are passed to
them.  The absence of side effects removes ordering constraints that could
restrict the freedom of the compiler to invoke pure functions out of order and
possibly in parallel.  Procedures and functions of this sort are also common in
pure functional languages like Haskell, and are exploited by compilers in order
to emit parallel code automatically due to their suitability for compiler-level
analysis.

Since pure procedures don't have side effects they are candidates for running on
accelerators in OpenCL.  Currently our prototype tool transforms pure procedures
to OpenCL kernels that \emph{do not} call other procedures, except for elemental
functions.

\subsection{New Procedures}

Borrowing ideas from ZPL, we introduce a concept of a region to Fortran with a
set of functions that allow programmers to work with subarrays in expressions.
In Fortran, these functions return a copy of an existing array or array section.
This is unlike ZPL, where regions are analogous to index sets and are used
primarily for address resolution within an array without dictating storage
related behavior.  The functions that we propose are similar in that they allow
a programmer to deal with index regions that are meaningful to their algorithm,
and automatically induce a halo (or ghost) cell pattern as needed in the
implementation generated by the compiler, where the size of an array is
increased to provide extra array elements surrounding the interior portion of
the array.  This is critical to reducing the amount of code that the programmer
is forced to write, as the halo cells are often where boundary errors and
off-by-one errors may occur in code that is manually generated.

Region functions are similar to the shift operator as they can be used to
reference portions of the array that are shifted with respect to the interior
portion.  However, unlike the shift operator, regions are not expressed in terms
of boundary conditions and thus don't explicitly \emph{require} a knowledge of
nor the application of boundary conditions locally.  Thus, as will be shown
below, regions are more suitable for usage by OpenCL thread groups which access
only local subsections of an array stored in global memory.

ForOpenCL provides two new procedures that are defined in Fortran and are used
in array-syntax operations.  Each procedure takes a integer array halo argument
that specifies the number of ghost cells on either side of a region, for each
dimension.  For example {\tt halo = [left, right, down, up]} specifies a halo
for a two-dimensional region.  These functions are:

\begin{itemize}

\item {\tt interior(array, halo)}: a pure function that returns a copy of
  the interior portion of the array specified by halo.

\item {\tt region(array, halo)}: a pure function that returns a copy of
  the portion of the array specified by halo.

\end{itemize}

It should be noted that the two functions {\tt interior} and {\tt region} are
pure and thus can be called from within a pure kernel procedure.  These two
functions are part of the language recognized by the compiler and though the two
functions return a copy of a portion of an array \emph{semantically}, the
compiler is not forced to actually make a copy and is free to enforce copy
semantics through other means.  In addition to these two new functions,
ForOpenCL provides the compiler directive, {\tt \!\$OFP KERNEL}, which specifies
that a pure subroutine can be transformed to an OpenCL kernel.

\subsection{Advantages}

There are several advantages to this style of programming using array
syntax, shifts, regions, and pure and elemental functions:

\begin{itemize}
\item There are no loops or index variables to keep track of.  Off by
  one index errors and improper handling of array boundaries are a
  common programming mistake.
\item The written code is closer to the algorithm, easier to
  understand, and is usually substantially shorter.
\item Semantically the intrinsic functions return arrays by value.
  This is usually what the algorithm requires.
\item Pure and elemental function are free from side effects, so it is
  easier for a compiler to schedule the work to be done in parallel.
\end{itemize}

Data parallelism has been called collection-oriented programming by
Blelloch~\cite{blelloch90}.  As the {\tt cshift} function and the array-valued
expressions all semantically return a value, this style of programming is also
similar to functional programming (or value-oriented programming).

Unfortunately, this style of programming has never really caught on because when
array syntax was first introduced in Fortran, performance of these features was
relatively poor and thus programmers shied away from using array syntax (even
recently, some are actively counseling against its usage because of performance
issues~\cite{Levesque:SC08}).  Thus the Fortran community was caught in a
classic ``chicken-and-egg'' conundrum: (1) programmers didn't use it because it
was slow; and (2) compilers vendors didn't improve it because programmers didn't
use it.  A goal of this paper is to demonstrate that parallel programs written
in this style of Fortran are maintainable and can achieve good performance on
accelerator architectures.

\subsection{Restrictions}

Only pure Fortran procedures are transformed into OpenCL kernels.  The
programmer must explicitly call these kernels from Fortran using the ForOpenCL
library described below.  It is also possible, using ROSE, to modify the calling
site so that the entire program can be transformed, but this functionality is
outside the scope of this paper.  Here we specifically examine transforming
Fortran procedures to OpenCL kernels.  Because OpenCL support is relatively new
to ROSE, some generated code must be modified.  For example, the {\tt
  \_\_global} attribute for kernel arguments was added by hand.

It is assumed that memory for all arrays reside on the device.  The programmer
must copy memory to and from the device.  In addition, array size (neglecting
ghost cell regions) must be multiples of the global OpenCL kernel size.

A kernel procedure (specified by the {\tt \!\$OFP KERNEL} directive) must be
pure and all array variables must be declared as contiguous.  A kernel procedure
may not call other procedures except for limited intrinsic functions (primarily
math), user-defined elemental functions, and the {\tt interior} and {\tt region}
functions.  Future work will address non-contiguous arrays (such as those that result
from strided access) by mapping array strides to gather/scatter-style memory
accessors.