### 2011-06-26 papers/GPUScA-2011/foropencl.tex
Version to ship to Matt.  Needs a couple of bib entries.

%
% I'm adding an outline here so I can see how everything fits together

%
% \title{Exploiting First-Class Arrays in Fortran for Accelerator Programming}
%

%
% \abstract
%

%
% \section{Introduction}
%    \subsection{Approach}
%    \subsection{Why Fortran?}
%    \subsection{Comparison to Other Languages}
%

%
% \section{Programming model}
%    \subsection{Fortran Syntax}   
%       \subsubsection*{Array notation}
%       \subsubsection*{Elemental functions}
%       \subsubsection*{Pure procedures}
%       \subsubsection*{Shift functions}
%       \subsubsection*{Regions}
%    \subsection{New Functions}
%    \subsection{Parallelism}
%    \subsection{Limitations}
%

%
% \section{Shallow Water Model}
%    \subsection{Equations}
%

%
% \section{Source-To-Source Transformations}
%    \subsection{ForOpenCL}
%       \subsubsection{array syntax}
%       \subsubsection{where construct}
%    \subsection{New functions}
%       \subsubsection{region}
%    \subsection{Static Analysis}
%       \subsubsection{Analysis not required}
%    \subsection{Simplifying Assumptions}
%

%
% \section{Performance}
%

%
% \section{Conclusions}
%

\documentclass{acm_proc_article-sp}

\usepackage{cite}
\usepackage{graphicx}
%\usepackage[cmex10]{amsmath}
%\usepackage{amssymb}
%\usepackage{array}
\usepackage{url}
\usepackage{listings}

\lstdefinestyle{FortranLike}{float,frame=lines,language=Fortran,commentstyle=\ttfamily,basicstyle=\ttfamily}
\lstdefinestyle{CLike}{float,frame=lines,language=C,commentstyle=\ttfamily,basicstyle=\ttfamily}
\lstdefinestyle{NoFloatCLike}{frame=lines,language=C,commentstyle=\ttfamily,basicstyle=\ttfamily}

\begin{document}

\title{ForOpenCL: Transformations Exploiting Array Syntax in Fortran for Accelerator Programming}

\numberofauthors{6}

\author{
% 1st. author
\alignauthor
Matthew J. Sottile\\
       \affaddr{Galois, Inc.}\\
       \affaddr{421 SW 6th Ave. Suite 300}\\
       \affaddr{Portland, OR 97204}\\
       \email{matt@galois.com}
% 2nd. author
\alignauthor
Craig E Rasmussen\\
       \affaddr{Los Alamos National Laboratory}\\
       \affaddr{CCS-7, MS B287}\\
       \affaddr{Los Alamos, NM 87545}\\
       \email{crasmussen@lanl.gov}
% 3rd. author
\alignauthor
Wayne N. Weseloh\\
       \affaddr{Los Alamos National Laboratory}\\
       %\affaddr{XTD-1, MS T082}\\
       %\affaddr{Los Alamos, NM 87545}\\
       \email{weseloh@lanl.gov}
\and  % use '\and' if you need 'another row' of author names
% 4th. author
\alignauthor
Robert W. Robey\\
       \affaddr{Los Alamos National Laboratory}\\
       %\affaddr{XCP-2, MS T086}\\
       %\affaddr{Los Alamos, NM 87545}\\
       \email{brobey@lanl.gov}
% 5th. author
\alignauthor
Daniel Quinlan\\
       \affaddr{Lawrence Livermore National Laboratory}\\
       \email{dquinlan@llnl.gov}
% 6th. author
\alignauthor
Jeffrey Overbey\\
       \affaddr{University of Illinois at Urbanna-Champaign}\\
       \email{overbey2@illinois.edu}
}

\maketitle

\begin{abstract}
  Emerging GPU architectures for high performance computing are well suited to a
  data-parallel programming model.  This paper presents preliminary work
  examining a programming methodology that provides Fortran programmers with access
  to these emerging systems.  We use array constructs in Fortran to
  show how this infrequently exploited, standardized language feature is easily
  transformed to lower-level accelerator code.  The transformations in ForOpenCL are based
  on a simple mapping from Fortran to OpenCL.  We demonstrate, using a
  stencil code solving the shallow-water fluid equations, that the performance
  of the ForOpenCL compiler-generated transformations is comparable with that of
  hand-optimized OpenCL code.
\end{abstract}

% A category with the (minimum) three required fields
\category{D.3.3}{Language Constructs and Features}{Concurrent programming structures}

\input{intro}
\input{programming-model}
\input{shallow-water}
\input{source-to-source}
%\input{static-analysis}
\input{performance}

\section{Conclusions}

The sheer complexity of programming for clusters of many or multi-core
processors with tens of millions threads of execution makes the simplicity of
the data-parallel model attractive.  The increasing complexity of
todays applications (especially in light of the increasing complexity
of the hardware) and the need for portability across architectures
make a higher-level and simpler programming model like data-parallel
attractive.

The goal of this work has been to exploit source-to-source transformations that
allow programmers to develop and maintain programs at a high-level of
abstraction, without coding to a specific hardware architecture.
Furthermore these transformations allow multiple hardware architectures
to be targeted without changing the high-level source.  It also removes the
necessity for application programmers to understand details of the accelerator
architecture or to know OpenCL.

%ACKNOWLEDGMENTS are optional
\section{Acknowledgments}
This research was supported by the Departement of Energy, Office of Science.

%%TODO \cite{chamberlain04zpl, roth97stencils}

\bibliographystyle{abbrv}
\bibliography{foropencl}

\balancecolumns
\end{document}