### 2011-10-27 papers/LOPe/intro.tex
Initial version.

It has been hard historically to create parallel programs.  The responsibility
(and difficulty) of creating \emph{correct} parallel programs can be viewed as
being spread between the programmer and the compiler.
Ideally we would like to have parallel languages that make it easy for the
programmer to express correct parallel programs (and conversely it should be difficult to express
incorrect parallel programs).  Unfortunately, many current languages and
standards place all of the responsibility on the user.  The best example of this
are programs written using MPI (Message Passing Interface), where the programmer
expresses all parallelism in terms of calls to library routines and the serial C or Fortran
compiler knows nothing about the parallel semantics of the combined language
plus MPI library.

For the purpose of discussion, we define a degree of difficulty term called the
joint Parallel Complexity Product (PCP),

    PCP = code\_complexity X compiler\_complexity

and suggest that PCP is roughly a constant over time, say the level of
complexity of writing an MPI program today, MPI\_PCP.

Of course one would hope that over time code complexity goes down as better
languages allow compilers to take on a greater share of the parallel complexity
burden.  This is somewhat true of UPC and Coarray Fortran.  With the PGAS
extensions, C and Fortran compilers are now aware of parallelism and now
generate message passing code that had been handled by the MPI library.  In some
instances the compiler is able to perform optimizations that it is not able to
do with a library based scheme like MPI \cite{CrayPGASPaper}

%%% Would like to say something here about HPC as a language based solution to reducing code complexity

However, in large part these languages are largely syntatic sugar for message
passing and do not provide a substatial decrease in code complexity
\cite{CAF_MPI_PGAS_COMPARISON}.  Fortunately, skilled programmers in the HPC
community have become accustomed to the level of complexity in an MPI program
and are welcome to considering the simplifications of HPC and Coarray Fortran.

The problem for programmers is that hardware is changing in ways that increase
the level of on-chip parallelism.  For ultimate performance, programmers must now
account for huge new levels of on-node parallelism at the same time they account
for off-node parallelism.  Thus PCP has suddenly increased with PCP = PCP\_Multi\_Core >> PCP\_MPI.
Since languages have not evolved that allow the compiler to take up this increase, unfortunately the
complexity for a programmer has dramatically increased.

A reasonable solution given todays language limitations is to use MPI (or a PGAS
language) \emph{plus} OpenMP for on-chip parallelism.  This is the solution proposed
by Cray and PGI \cite{BOF_SC10} (currently Chapel does not target GPUs\cite{Brad?}).
Other choices for expressing on-chip parallelism are OpenCL\cite{OPENCL} and NVIDIA's CUDA.

In this paper we examine a language-based paradigm that allows the compiler to
take on a larger portion of the PCP burden.  Anyone programming in OpenCL or
CUDA is aware that explicit loop structures over array elements in a serial
program are removed and replaced by a kernel program that is run over all of
the