### 2011-11-16 papers/IJCSE/rose-transform.tex
starting new subsection

\subsection{Kernel transformation}

In addition to the transformation of the code at call sites to copy
data back and forth from the device and invoke the OpenCL kernel, we
must also provide transformations of the Fortran kernel logic to the
equivalent OpenCL C code.  This transformation is fairly
straightforward, and maps a restricted subset of Fortran directly to
its C equivalent.  The transformation is based on a regular traversal
of the ROSE Sage AST that represents the Fortran code.  The visitor
used in the traversal handles five types of Sage AST Nodes: {\tt
  SgAllocateStatment}, {\tt SgFunctionDeclaration},
{\tt SgVariableDeclaration}, {\tt SgFunctionCallExp}, and
{\tt SgExprStatement}.  These encompass the primary activities that are
commonly encountered in computational kernels: allocation, function invocation,
variable and function declaration, and expressions that correspond to
arithmetic operations.