### 2011-11-29 papers/IJCSE/abstract.tex
fixedits

\begin{abstract}
  Emerging GPU architectures for high performance computing are well
  suited to a data-parallel programming model.  This paper presents
  preliminary work examining a programming methodology that provides
  Fortran programmers with access to these emerging systems.  We use
  array constructs in Fortran to show how this infrequently exploited,
  standardized language feature is easily transformed to lower-level
  accelerator code.  The transformations in ForOpenCL are based on a
  simple mapping from Fortran to OpenCL.  We demonstrate, using a
  stencil code solving the shallow-water fluid equations, that the
  performance of the ForOpenCL compiler-generated transformations is
  comparable with that of hand-optimized OpenCL code.
\end{abstract}