### 2011-11-29 papers/IJCSE/foropencl.tex

%%
\documentclass{doublecol-new}
%%

%\usepackage{natbib, stfloats}
\usepackage{cite}
\usepackage{graphicx}
%\usepackage[cmex10]{amsmath}
%\usepackage{amssymb}
%\usepackage{array}
\usepackage{url}
\usepackage{listings}

\lstdefinestyle{FortranLike}{float,frame=lines,language=Fortran,commentstyle=\ttfamily,basicstyle=\ttfamily}
\lstdefinestyle{CLike}{float,frame=lines,language=C,commentstyle=\ttfamily,basicstyle=\ttfamily}
\lstdefinestyle{NoFloatCLike}{frame=lines,language=C,commentstyle=\ttfamily,basicstyle=\ttfamily}

\begin{document}

\thispagestyle{plain}

\LRH{M. Sottile et~al.}

\RRH{ForOpenCL: Transformations Exploiting Array Syntax in Fortran for Accelerator Programming}

\title{ForOpenCL: Transformations Exploiting Array Syntax in Fortran for Accelerator Programming}

% 1st. author
\authorA{Matthew J. Sottile}
\affA{Galois, Inc.\\
% Portland, OR 97204\\
 E-mail: mjsottile@computer.org}

% 2nd. author
\authorB{Craig E Rasmussen}
\affB{Los Alamos National Laboratory\\
% Los Alamos, NM 87545\\
 E-mail: crasmussen@lanl.gov}

% 3rd. author
\authorC{Wayne N. Weseloh}
\affC{Los Alamos National Laboratory\\
% Los Alamos, NM 87545\\
 E-mail: weseloh@lanl.gov}

% 4th. author
\authorD{Robert W. Robey}
\affD{Los Alamos National Laboratory\\
% Los Alamos, NM 87545\\
 E-mail: brobey@lanl.gov}

% 5th. author
\authorE{Daniel Quinlan}
\affE{Lawrence Livermore National Laboratory\\
% Livermore, CA 94550\\
 E-Mail: dquinlan@llnl.gov}

% 6th. author
\authorF{Jeffrey Overbey}
\affF{University of Illinois at Urbanna-Champaign\\
% Champaign, IL 61820\\
 E-mail: overbey2@illinois.edu}

%% bio
\begin{bio}
M. Sottile completed his PhD in 2006 from the University of New Mexico
from the Department of Electrical and Computer Engineering.  He is
presently working as a researcher at Galois, Inc. in Portland, Oregon
leading research work efforts in scientific and high performance computing.
His research interests include programming languages, compilation techniques,
computational physics and applied mathematics.

\noindent Craig has a bio.

\noindent Wayne has a bio.

\noindent Bob has a bio.

\noindent Dan has a bio.

\noindent Jeff has a bio.
\end{bio}

\maketitle

\input{abstract}

% A category with the (minimum) three required fields
%\category{D.3.3}{Language Constructs and Features}{Concurrent programming structures}

\input{intro}
\input{programming-model}
\input{shallow-water}
\input{source-to-source}
%\input{static-analysis}
\input{performance}

\section{Conclusions}

The sheer complexity of programming for clusters of many or multi-core
processors with tens of millions threads of execution makes the simplicity of
the data-parallel model attractive.  The increasing complexity of
today's applications (especially in light of the increasing complexity
of the hardware) and the need for portability across architectures
make a higher-level and simpler programming model like data-parallel
attractive.

The goal of this work has been to exploit source-to-source transformations that
allow programmers to develop and maintain programs at a high-level of
abstraction, without coding to a specific hardware architecture.
Furthermore these transformations allow multiple hardware architectures
to be targeted without changing the high-level source.  It also removes the
necessity for application programmers to understand details of the accelerator
architecture or to know OpenCL.

%ACKNOWLEDGMENTS are optional
\section{Acknowledgments}
This work was supported in part by the Department of Energy Office of Science,
Advanced Scientific Computing Research.

%%TODO \cite{chamberlain04zpl, roth97stencils}

%\bibliographystyle{abbrv}
\bibliographystyle{plain}
\bibliography{foropencl}

\end{document}