### 2011-12-11 papers/IJCSE/intro.tex
edits to intro

\section{Introduction}
\label{sec:intro}

This paper presents a compiler-level approach for targeting a single
program to multiple, and possibly fundamentally different, processor
architectures.  This technique allows the application programmer to
adopt a single, high-level programming model without sacrificing
performance.  We suggest that existing data-parallel features in
Fortran are well-suited to applying automatic transformations that
generate code specifically tuned for different hardware architectures
using low-level programming models such as OpenCL.  For algorithms
that can be easily expressed in terms of whole array, data-parallel
operations, writing code in Fortran and transforming it automatically
to specific low-level implementations removes the burden of creating
and maintaining multiple versions of architecture specific code.

The peak performance of these newer accelerator architectures can be
substantial.  Intel expects a teraflop for the SGEMM benchmark with
their Knights Ferry processor while the performance of the M2090
NVIDIA Tesla processor is in the same
neighborhood~\citep{hpcwire11manycore}.  Unfortunately the performance
that many of the new accelerator architectures offer comes at a cost.
Architectural changes are trending toward multiple heterogeneous cores
and less of a reliance on superscalar instruction level parallelism
and hardware managed memory hierarchies (such as traditional caches).

These changes place a heavy burden on application programmers as they
work to adapt to these new systems.  An especially challenging problem
is not only how to program to these new architectures --- considering
the massive scale of concurrency available --- but also how to design
programs that are portable across the changing landscape of computer
architectures.  How does a programmer write one program that can
perform well on both a conventional multicore CPU \emph{and} a GPU (or
any other emerging many-core architectures)?

A directive-based approach, such as OpenMP or the Accelerator
programming model from the Portland Group~\citep{pgi10accelerator} and
now generalized in the OpenACC standard~\citep{openacc11std}, is one
solution to this problem.  However, in this paper we take a somewhat
different approach not requiring explicit declaration of parallelism,
instead relying on implicitly parallel subsets of the Fortran
language.  A common theme amongst the new processors is the emphasis
on data-parallel programming.  This model is well-suited to
architectures that are based on either vector processing or massively
parallel collections of simple cores.  The recent CUDA and OpenCL
programming languages are intended to support this programming model.

The problem with OpenCL and CUDA is that they expose too much detail
about the machine architecture to the programmer~\citep{wolfe08gpgpu}.
The programmer is responsible for explicitly managing memory
(including the staging of data back and forth between the host CPU and
the accelerator device) and specifically taking into account
architectural differences (such as whether the architecture contains
vector units).  While these languages have been attractive as a method
for early adopters to utilize these new architectures, they are less
attractive to programmers who do not have the time or resources to
manually port their code to every new architecture and programming
model that emerges.

\subsection{Approach}

We demonstrate that a subset of Fortran maps surprisingly well onto
GPUs when transformed to OpenCL kernels.  This data-parallel subset
includes: array syntax using assignment statements and binary
operators, array constructs like {\tt WHERE}, and the use of pure and
elemental functions.  In addition, we provide new functions that
explicitly take advantage of the stencil geometry of the problem
domain we consider.  Note that this subset of the Fortran language is
implicitly parallel.  This programming model \emph{does not require
  explicit declaration of parallelism within the program.}  In
addition, programs are expressed using entirely standard Fortran so it
can be compiled for and executed on a single core without concurrency.

Transformations are supplied that provide a mechanism for converting
Fortran procedures written in the Fortran subset described in this
paper to OpenCL kernels.  We use the ROSE compiler
infrastructure~\citep{rose-web} to develop these transformations.
ROSE uses the Open Fortran Parser~\citep{ofp:sf} to parse Fortran 2008
syntax and can generate C-based OpenCL.  Since ROSE's intermediate
representation (IR) was constructed to represent multiple languages,
it is relatively straightforward to transform high-level Fortran IR
nodes to C OpenCL nodes.  This work is also applicable to
transformations to vendor-specific languages, similar to OpenCL, such
as the NVIDIA CUDA language.

Transformations for arbitrary Fortran procedures are not attempted.
Furthermore, a mechanism to transform the calling site to
automatically invoke OpenCL kernels is not provided at this time.
While it is possible to accomplish this task within ROSE, it is
considered outside the scope of this paper.  However, ForOpenCL
provides via Fortran interfaces a mechanism to call the C OpenCL
runtime and enable Fortran programmers to access OpenCL kernels
generated by the supplied transformations.

We study the automatic transformations for an application example that
is typical of stencil codes that update array elements according to a
fixed pattern.  Stencil codes are often employed in applications based
on finite-difference or finite-volume methods in computational fluid
dynamics (CFD).  The example described later in this paper is a simple
shallow-water model in two dimensions using finite volume methods.
Stencil-like patterns appear in a number of other contexts as well.
In image processing, they appear in convolution-based algorithms in
which small kernels are convolved with an image to implement
denoising, edge detection, and other common operators.  Similar
stencil operators appear in general signal processing applications as
well.
%% more?

Finally, we examine the performance of the Fortran data-parallel abstraction
when transformed to OpenCL to run on GPU architectures.  The performance of
automatically transformed code is compared with a hand-optimized OpenCL
version of the shallow-water code.

We do not perform any additional analysis of the code to identify
parallelism beyond that present in the data parallel operations that we
focus on in this paper.  Additional program
analysis methods may be investigated to study their applicability
in future versions of this work.