### 2012-03-13 papers/IJCSE/foropencl.tex
edits for final version

%%
\documentclass{doublecol-new}
%%

\usepackage{natbib, stfloats}
\usepackage{cite}
\usepackage{graphicx}
%\usepackage[cmex10]{amsmath}
%\usepackage{amssymb}
%\usepackage{array}
\usepackage{url}
\usepackage{listings}

\lstdefinestyle{FortranLike}{float,frame=lines,language=Fortran,commentstyle=\ttfamily,basicstyle=\ttfamily}
\lstdefinestyle{CLike}{float,frame=lines,language=C,commentstyle=\ttfamily,basicstyle=\ttfamily}
\lstdefinestyle{NoFloatCLike}{frame=lines,language=C,commentstyle=\ttfamily,basicstyle=\ttfamily}

\begin{document}

\thispagestyle{plain}

%% ANONYMIZED
\LRH{M. Sottile et~al.}
%\LRH{ }

\RRH{ForOpenCL: Transformations Exploiting Array Syntax in Fortran for Accelerator Programming}

\title{ForOpenCL: Transformations Exploiting Array Syntax in Fortran for Accelerator Programming}

%% ANONYMIZED
%\authorA{Author names removed for review}
% 1st. author
\authorA{Matthew J. Sottile}
\affA{Galois, Inc.\\
 Portland, OR 97204\\
 E-mail: mjsottile@computer.org}

% 2nd. author
\authorB{Craig E Rasmussen}
\affB{Los Alamos National Laboratory\\
 Los Alamos, NM 87545\\
 E-mail: crasmussen@lanl.gov}

% 3rd. author
\authorC{Wayne N. Weseloh}
\affC{Los Alamos National Laboratory\\
 Los Alamos, NM 87545\\
 E-mail: weseloh@lanl.gov}

% 4th. author
\authorD{Robert W. Robey}
\affD{Los Alamos National Laboratory\\
 Los Alamos, NM 87545\\
 E-mail: brobey@lanl.gov}

% 5th. author
\authorE{Daniel Quinlan}
\affE{Lawrence Livermore National Laboratory\\
 Livermore, CA 94550\\
 E-Mail: dquinlan@llnl.gov}

% 6th. author
\authorF{Jeffrey Overbey}
\affF{National Center for Supercomputing Applications\\
 Urbana, IL 61801\\
 E-mail: overbey2@illinois.edu}

\input{abstract}

\KEYWORD{Fortran; OpenCL; Accelerator programming; GPU programming; code generation.}

%% bio
%\begin{bio}
%M. Sottile completed his PhD in 2006 from the University of New Mexico
%from the Department of Electrical and Computer Engineering.  He is
%presently working as a researcher at Galois, Inc. in Portland, Oregon
%leading research work efforts in scientific and high performance computing.
%His research interests include programming languages, compilation techniques,
%computational physics and applied mathematics.

%\noindent Craig has a bio.

%\noindent Wayne has a bio.

%\noindent Bob has a bio.

%\noindent Dan has a bio.

%\noindent Jeff has a bio.
%\end{bio}

\maketitle

% A category with the (minimum) three required fields
%\category{D.3.3}{Language Constructs and Features}{Concurrent programming structures}

\input{intro}
\input{programming-model}
\input{shallow-water}
\input{source-to-source}
%\input{static-analysis}
\input{performance}

\section{Conclusions}

The sheer complexity of programming for clusters of many or multi-core
processors with tens of millions threads of execution makes the simplicity of
the data-parallel model attractive.  The increasing complexity of
today's applications (especially in light of the increasing complexity
of the hardware) and the need for portability across architectures
make a higher-level and simpler programming model like the data-parallel
model attractive.

The goal of this work has been to exploit source-to-source transformations that
allow programmers to develop and maintain programs at a high-level of
abstraction, without coding to a specific hardware architecture.
Furthermore these transformations allow multiple hardware architectures
to be targeted without changing the high-level source.  It also removes the
necessity for application programmers to understand details of the accelerator
architecture or to know OpenCL.

%ACKNOWLEDGMENTS are optional

\section{Acknowledgments}
This work was supported in part by the Department of Energy Office of Science,
Advanced Scientific Computing Research.

%%TODO \cite{chamberlain04zpl, roth97stencils}

%\bibliographystyle{abbrv}
\newcommand{\newblock}{}
\bibliographystyle{plainnat}
\bibliography{foropencl}

%% this was only required for review purposes - remove in final
%\newpage
%
%\input{note-to-reviewers}

\end{document}