### 2015-01-22 papers/LOPe/local.tex
Starting work on EuroPar version.

\documentclass{acm_proc_article-sp}

\usepackage{cite}
\usepackage{graphicx}
%\usepackage[cmex10]{amsmath}
%\usepackage{amssymb}
%\usepackage{array}
\usepackage{url}
\usepackage{listings}

\lstdefinestyle{FortranLike}{float,frame=lines,language=Fortran,commentstyle=\ttfamily,basicstyle=\ttfamily}
\lstdefinestyle{CLike}{float,frame=lines,language=C,commentstyle=\ttfamily,basicstyle=\ttfamily}
\lstdefinestyle{NoFloatCLike}{frame=lines,language=C,commentstyle=\ttfamily,basicstyle=\ttfamily}

\begin{document}

\title{Locally-Oriented Programming: A simple programming model for stencil computations}

\numberofauthors{4}

\author{
% 1st. author
\alignauthor
Craig E Rasmussen\\
       \affaddr{Los Alamos National Laboratory}\\
%       \affaddr{CCS-7, MS B287}\\
       \affaddr{Los Alamos, NM 87545}\\
       \email{crasmussen@lanl.gov}
% 2nd. author
\alignauthor
Matthew J. Sottile\\
       \affaddr{Galois, Inc.}\\
%       \affaddr{421 SW 6th Ave. Suite 300}\\
       \affaddr{Portland, OR 97204}\\
       \email{mjsottile@computer.org}
% 3rd. author
\alignauthor
Dan Nagle\\
       \email{dannagle@verizon.net}
% 4th. author
\alignauthor
Soren Rasmussen\\
       \affaddr{Utah State University}\\
       \email{soren.rasmussen@aggiemail.usu.edu}
}

\maketitle

\begin{abstract}
Some stuff...

%%Emerging GPU architectures for high performance computing are well suited to a data-parallel programming model.  However, programming for these architectures currently requires learning a new language, e.g., OpenCL or CUDA, or employing OpenMP compiler directives.  This paper presents preliminary work examining how programming with a local orientation can be employed in Fortran (or using a special array library in C++) to provide simple access GPU architectures.  A locally-oriented programming model is especially useful for stencil-based codes or algorithms requiring the application of convolution kernels.  In this programming model, a programmer codes the algorithm {\it only} for a single array element, but has read-only access to a small sub-array surrounding the given array element, so that a stencil can be applied.  We demonstrate how a locally-oriented programming model can be adopted using source-to-source program transformations in conjuction with a compiler-based infrastructure like ROSE.
\end{abstract}

% A category with the (minimum) three required fields
%\category{D.3.3}{Language Constructs and Features}{Concurrent programming structures}

%\input{intro}
%\input{motivation}
\input{programming-model}
%\input{program-restructuring}
%\input{static-analysis}
%\input{performance}

\section{Conclusions}

Some stuff...

%%The sheer complexity of programming for clusters of many or multi-core processors with tens of millions threads of execution makes the simplicity of the data-parallel model attractive.  The increasing complexity of today's applications (especially in light of the increasing complexity of the hardware) and the need for portability across architectures make a higher-level and simpler programming model like data-parallel attractive.

%%The goal of this work has been to exploit source-to-source transformations that allow programmers to develop and maintain programs at a high-level of abstraction, without coding to a specific hardware architecture.  Furthermore these transformations allow multiple hardware architectures to be targeted without changing the high-level source.  It also removes the necessity for application programmers to understand details of the accelerator architecture or to know OpenCL.

%ACKNOWLEDGMENTS are optional
\section{Acknowledgments}
This work was supported in part by the Department of Energy Office of Science,
Advanced Scientific Computing Research.

%%TODO \cite{chamberlain04zpl, roth97stencils}

%\bibliographystyle{abbrv}
%\bibliography{local}

\balancecolumns
\end{document}