### 2015-01-24 papers/LOPe/local.tex
Working on paper layout and program models section primarily.

%%%%%%%%%%%%%%%%%%%%%%% file typeinst.tex %%%%%%%%%%%%%%%%%%%%%%%%%
%
% This is the LaTeX source for the instructions to authors using
% the LaTeX document class 'llncs.cls' for contributions to
% the Lecture Notes in Computer Sciences series.
% http://www.springer.com/lncs       Springer Heidelberg 2006/05/04
%
% It may be used as a template for your own input - copy it
% to a new file with a new name and use it as the basis
% for your article.
%
% NB: the document class 'llncs' has its own and detailed documentation, see
% ftp://ftp.springer.de/data/pubftp/pub/tex/latex/llncs/latex2e/llncsdoc.pdf
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass[runningheads,a4paper]{llncs}

\usepackage{amssymb}
\setcounter{tocdepth}{3}
\usepackage{graphicx}

\usepackage{url}
\urldef{\mailsa}\path|rasmus@uoregon.edu, mjsottile@computer.org,|
\urldef{\mailsb}\path|dannagle@verizon.net, soren.rasmussen@aggiemail.usu.edu|    
\newcommand{\keywords}[1]{\par\addvspace\baselineskip
\noindent\keywordname\enspace\ignorespaces#1}

\begin{document}

\mainmatter  % start of an individual contribution

% first the title is needed
\title{Locally-Oriented Programming:\\ A Simple Programming Model for Stencil-Based Computations
       on Multi-Level Distributed Memory Architectures}

% a short form should be given in case it is too long for the running head
\titlerunning{Locally-Oriented Programming}

\author{Craig Rasmussen%
\thanks{The authors would like to thank Robert Robey and Wayne Weseloh
at Los Alamos National Laborabory for several stimulating
conversations with respect to programming models.}%
\and Matthew Sottile\and Dan Nagle\and Soren Rasmussen}

\authorrunning{Locally-Oriented Programming}
% (feature abused for this document to repeat the title also on left hand pages)

% the affiliations are given next; don't give your e-mail address
% unless you accept that it will be published
\institute{441 McKenzie Hall,\\
5246 University of Oregon, Eugene, Oregon, USA\\
\mailsa\\
\mailsb\\
\url{https://openfortran.uoregon.edu}}

%
% NB: a more complex sample for affiliations and the mapping to the
% corresponding authors can be found in the file "llncs.dem"
% (search for the string "\mainmatter" where a contribution starts).
% "llncs.dem" accompanies the document class "llncs.cls".
%

%\toctitle{Lecture Notes in Computer Science}
%\tocauthor{Authors' Instructions}
\maketitle

\begin{abstract}

Emerging hybrid accelerator architectures for high performance computing are often well suited for
usage of a data-parallel programming model.  However, programmers for these emerging architectures
face a steep learning curve that frequently requires learning a new language, e.g., OpenCL or CUDA,
or employing OpenMP compiler directives.  Furthermore, the distributed (and frequently multi-level)
nature of the memory organization of clusters of these architectures provides an additional level of
complexity for the programmer.  This paper presents preliminary work examining how programming with
a local orientation can be employed to provide simple access to accelerator architectures.  A
locally-oriented programming model is especially useful for stencil-based codes or algorithms
requiring the application of convolution kernels.  In this programming model, a programmer codes the
algorithm by modifying \emph{only a single array element}, but has read-only access to a small
sub-array surrounding the given array element, so that a stencil can be applied.  We demonstrate how
a locally-oriented programming model can be adopted as an embedded, domain-specific language using
source-to-source program transformations.

\keywords{domain specific language, stencil compilers, distributed memory parallelism}
\end{abstract}

\input{motivation}
\input{programming-model}
\input{appendix-a}

\subsection{Acknowledgments}
This work was supported in part by the Department of Energy Office of Science,
Advanced Scientific Computing Research.

\end{document}