### 2015-01-24 papers/LOPe/appendix-a.tex
Initial version.

\section*{Appendix: A Fortran Primer}

This section provides a brief description of Fortran syntax for readers unfamiliar with Fortran syntax.