### 2015-01-27 papers/LOPe/appendix-a.tex
Finished second reading of programming model.

\section*{Appendix: A Fortran Primer for the C Programmer}

This section provides a brief description of Fortran syntax for readers unfamiliar with Fortran syntax.