### 2015-01-27 papers/LOPe/coarray-extensions.tex

\section{Coarray Fortran Extensions}

%%We highlight the Fortran elemental source-code abstraction because it provides a local orientation similar to that of OpenCL or CUDA kernel functions that were designed to exploit streaming, highly-threaded architectures like GPUs.  Elemental functions are pure in that they are guaranteed to be free of side effects such as I/O.  Consider a convolution where a 3x3 filter is applied to individual pixel elements in a photograph to average of blur the original photograph.  A convolution is similar to stencil operations in computational fluid dynamics that are used to obtain spatial derivatives of state variables.  Simple code is shown with extentions to Fortran shown in capital letters.  The extensions could be provided to C or Fortran functions via compiler directives rather than explicit language syntax.

Consider the \texttt{Laplace} concurrent function discussed above.  In this section we demonstrate
how this function can be called in a context that allows full access to all of the interior elements
of the array, as well as array elements within the logically exterior, halo boundary region.  Topics
highlighted in this section are:
1. Distributed memory array allocation;
2. Explicit memory placement;
3. Remote memory transfer; and
4. Remote execution.
This description is within the context of a Domain Specific Language (DSL), again extending Fortran
2008.  As shorthand, these specific extensions are referred to as CAFe, for Coarray Fortran extensions.

\subsection{CAFe Example}

We start with the declaration of an array (to be allocated later),
\begin{verbatim}
  real, allocatable, HALO(1:*:1,1:*:1),          &
        dimension(:,:), codimensions(:,:) :: U
\end{verbatim}
with an explicit halo size and with two local dimensions (rank) and two distributed memory
codimensions (corank).  The corank of the array is chosen to be identical to the rank of
the array so that the logical process topology aligns is a way that allows a natural
halo exchange between logically neighboring processes.  For example, if my process
location is \texttt{[pcol,prow]}, then the right-hand halo for my local array \texttt{U}
can be obtained by the assignment \texttt{U(M+1,:) = U(1,:)[pcol+1,prow]} where the size
and cosize of \texttt{U} are given by the allocation statement,
\begin{verbatim}
  allocate( U(0:M+1,0:N+1)[MP,*] )
\end{verbatim}

This allocation statement specifies (given the one element halo size provided earlier for
\texttt{U}) that the left halo column is \texttt{U(0,:)}, the right column is \texttt{U(M+1,:)}, the
bottom row is \texttt{U(:,0)} and the top row is \texttt{U(:,N+1)}, leaving the enterior region
\texttt{U(1:M,1:N)}.  In this allocation statement, the total number of process rows \texttt{NP} can
be obtained at runtime, but may not be explicitly provided (according to CAF rules) because the
actual number of participating processes (in Fortran called images) is variable, depending on how
many processes are requested at startup.  In this discussion, it is \emph{assumed} that there are no
holes in the logical processor topology, thus $MP*NP = P$, where $P$ is the total number of
participating processes (images).

So far --- except for the \texttt{HALO} attribute --- this is totally standard Fortran (2008).  For CAFe,
we extend the concept of Fortran images to be hierachical.  By this we mean that each image \emph{may}
have a subimage (or subimages), but this subimage is not visible to other regular Fortran images.
Subimages also execute differently than normal images.  Subimages are task based while images all
execute a single program but with different (multiple) data (SPMD).  A task can be given to a subimage,
but execution on the subimage terminates once the task is finished.  Memory on a subimage is
permanent, however, and must be specifically allocated and deallocated.

One obtains a subimage by executing a function call,
\begin{verbatim}
   device = get_subimage(1)
\end{verbatim}
where the integer argument representing one of attached hardware devices.  If the function
fails (e.g., the requested device is unavailable) it returns the image number,
\texttt{this\_image()}, of the process that is executing the current program.  Returning
the current image allows program execution to proceed correctly even if there are no
attached devices.

Once a subimage is obtained, memory on the device may be allocated,
\begin{verbatim}
  if (device /= this_image()) then
    allocate(U[device], HALO_SRC=U)   [[device]]
  end if
\end{verbatim}
There are four points to note regarding this memory allocation: 1. Memory is only allocated if a
subimage has been obtained; 2. The location where memory is allocated is denoted by regular coarray
notation \texttt{U[device]}; 3. The allocated size and halo attribute of the new array are assumed
from the previously allocated local array \texttt{U} via the notation \texttt{HALO\_SRC=U} (using
\texttt{HALO\_SRC} will also initially copy \texttt{U} to the subimage); and finally 4. The
allocation itself is \emph{executed} on the subimage device with the notation \texttt{[[device]]}.

Fortran uses square bracket notation \texttt{[image]} to specify on what process the
memory reference is physically located.  Square brackets are a visual clue to the
programmer that the memory reference may be remote and therefore potentially suffer a
performance penalty.  CAFe extends this by employing double-bracket notation to indicate
possibly \emph{remote subimage execution}.

Execution of the \texttt{Laplace} task (defined previously) is done using the Fortran \texttt{do}
\texttt{concurrent} construct as shown,
\begin{verbatim}
   do while (.not. converged)
      do concurrent (i=1:M, j=1:N)   [[device]]
         call Laplace( U(i,j)[device] )
      end do
      call HALO_TRANSFER(U, BC=CYCLIC)
   end do
\end{verbatim}
There are several points that require highlighting: 1. Iteration occurs over the interior
of the array domain, \texttt{(i=1:M, j=1:N)}; 2. Execution of the loop body occurs on the
specific subimage indicated by \texttt{[[device]]}; 3. Execution of the iterates may occur
in any order, even \emph{concurrently}; 4. The local element of the array (as
defined above in reference to the definition of the concurrent procedure
\texttt{Laplace}) is given by the indices \texttt{(i,j)}; 5. Location of memory for the
task is to be taken from the subimage as noted by \texttt{[device]}; 5. All threads must finish
execution of the loop body before further execution of the program proceeds; and 6. Transfer of
all requisite halo regions is effected by the call to the new LOPe intrinsic function
\texttt{HALO\_TRANSFER()}.  This function is a synchronization event in that all images must
complete the halo transfer before program execution continues.

Note that a transfer of halo memory is necessary after each completion of the do concurrent loop.
This must be done in order for the halo region of a coarray on a given process to be consistent
with the corresponding interior of the coarray on a logically neighboring process.

%% However, recall that with the LOPe programming model only the local element is mutable; thus, the halo region of an array needs update.

Finally, memory for the entire array \texttt{U} can be copied from the subimage device
with the statement,
\begin{verbatim}
  U = U[device]
\end{verbatim}
add memory deallocation is similar to memory allocation:
\begin{verbatim}
   deallocate(U)
   if (device /= this_image()) then
      deallocate(U[device])   [[device]]
   end if
\end{verbatim}

%%\subsection{Execution Symantics and Memory Management}

%%This section describes the hierarchical memory layout and how memory consistency between the 3-levels of memory is maintained.  It describes what is the compiler's responsibility and what is the programmer's responsibility.

%% Completion of a do concurrent construct indicates that all executing threads (if a threading model is used by the compiler) have completed and that all memory on the executing subimage is in a consistent state.

%%The {\tt HALO} function returns a copy of the 3x3 region of {\tt a} and its surrounding neighbors.  The {\tt convolve} function is free of race conditions because of the copy semantics of {\tt HALO} and the implied synchronization of the {\tt CONCURRENT} attribute, whereby no output variables can be updated before all threads have completed execution.  In addition, while any thread may load from an extended region about \emph{its} element with the {\tt HALO} function, it may only store into its own elemental location.

%% A note from a conversation with Matt regarding memory management
%%

%The Fortran language has much tighter restrictions on aliasing than does C.
%So unless a variable has the pointer or target attribute, it cannot be aliased.
%Thus the compiler is able to agressively optimized for memory movement between
%the CPU and accelerator.  However, because the design philosophy of coarrays
%is that memory transfer between images can be expensive, the programmer must
%explicitly transfer memory between images with explicit syntax with square
%bracket notation, i.e., $a[1] = a[2]$.  So we allow the compiler to manage
%memory within an image but require the use of {\tt halo\_exchange} for the
%transfer of halo memory between images.

%Fortran currently supports:

%1. Array syntax: e.g., C = A + B, where A, B, C are arrays.  Note that this is implicitly a loop structure, but that no loop indices need be provided.  Also the programmer need not specify where in memory these arrays reside.  Thus this high level syntax allows the compiler more freedom in both memory placement (even across distributed memory nodes) and in runtime code execution (individual array element may be computed by different hardward threads).  This is a simple example and it is a research question as to what compiler directives would be useful for memory placement and other directions to the compiler for efficient code generation.

%2. Pure procedures:  Fortran has syntax for specifying procedures that have no side effects during execution.  Specifying code that is side-effect free code is important information to provide to the compiler so that it can generate efficient multi-threaded code.

%3. Pure elemental procedures: Fortran has syntax for specifying procedures that take only scalar arguments, but may be applied across array elements.  Elemental procedures are ideal for writing code to be executed within a hardware thread.  They resemble OpenCL kernels, but are simpler because they leave all indexing up to the compiler.

%We have determined that additional syntax is needed, in addition to the three language features described above, to allow programmers the ability to express code in Fortran to be targeted for multi-threaded hardware architectures like GPUs.  This additional syntax is provided by functions that return a copy of a small region of memory surrounding an array element (as seen within an elemental procedure) and with functions for thread synchronization.  This additional syntax will allow pure procedures to perform stencil and other convolution-like operations on a copy of memory, synchronize, then store the computed results back to the array element associated with the given thread.

%In addition Fortran has syntax like the target attribute the specifies when variables can be aliased.  This allows for much easier program analysis as the compiler knows that ordinary variables cannot be aliased.  Fortran also has excellent facilities for interoperability with C so that programming in a mixed language invironment is easily accomplished, including interoperability with native Fortran arrays.

\input{coarray-comparison}

Note that the use of halo cells is the normal way that large and complex MPI and CAF
programs are implemented.  LOPe proposes to formalize this common pattern into the Fortran
language allowing the compiler access to this information in order to spread computation over more
hardware resources, improve performance, and to reduce complexity for the programmer.