### 2015-01-27 papers/LOPe/programming-model.tex

\section{Programming Model}

The LOPe programming model (for Locally-Oriented Programming extensions) restricts the programmer to
a local view of the index space of an array.  Within a LOPe function, only a single array element (called
the local element) is mutable.  In addition, a small halo region surrounding the local element is
visible to the programmer, but this region is immutable.  Restricting the programmer to a local
index space serves to reduce complexity by separating all boundary-value and processor-topology
concerns from the implementation of the algorithm.

%%This reduction in complexity reduces programming errors.  While developing the convolution example described later in the paper we made an indexing error in applying the 2D stencil loops in the standard serial Fortran test implementation.  This error required over 3 hours of programming time to repair.  First the error had to be isolated to the function implementing the convolution (it was first thought to be in the complicated tiff image output routine as the convolution code was ``thought'' to be too simple to wrong.  Then the index error has to be understood.  As will be seen, LOPe makes it more difficult to make these errors as Fortran array intrics can be used.  In some instance, the restricted semantics of LOPe allows the compiler to catch errors (e.g., some errors involving race conditions).

LOPe was implemented as an embedded domain specific language (DSL).  Fortran was chosen as the base
language for LOPe, although in principle, the same techniques could be applied to language such as C
or C++ (in conjuction with some form of array extensions such as those proposed by Intel).  Fortran
was chosen because it already provides a rich array-based syntax.  Readers who are unfamiliar with
Fortran syntax may wish to consult Appendix A, for a brief description of Fortran notation.

\subsection{LOPe Syntax Extensions}

There are only a few syntax additions required for a LOPe program.  These additions include syntax for
describing halo regions and concurrent procedures.  In all of the code examples that follow, DSL
syntax additions are highlighted by the usage of capitalization for keywords that are either new or
that acquire new usage.

\textbf{Halo regions.}
The principle semantic element of LOPe is the concept of a halo region.  A halo is an ``artificial''
or ``virtual'' region surrounding an array that contains boundary-value information.  Halo (also
called ghost-cell) regions are commonly employed to unify array indexing schemes in the vicinity of
an array boundary, so that an array may be referenced using indices that fall ``outside'' of the
logical domain of the array.  In LOPe, the halo region is given explicit syntax so that the compiler
can exploit this knowledge.  For example, a halo region can be declared with a statement of the form,
\begin{verbatim}
  real, allocatable, dimension(:), HALO(1:*:1) :: A
\end{verbatim}
This statement indicates that \texttt{A} is a rank one array, will be allocated later (when it will
be provided with a specific size), and has a halo region of one element surrounding the array on
either side.  Notation such as \texttt{M:*:N} is to be read as describing a halo region of
\texttt{M} elements to the left, \texttt{N} elements to the right, and with an arbitrary number of
elements in the middle (denoted by the $*$ symbol).  When used to describe a formal parameter of a
function, the halo size may be ``assumed'' (denoted by a ``:'' symbol substituted in place of an
explicit size) as is the following,
\begin{verbatim}
  real, HALO(:,:) :: U
\end{verbatim}
which in this instance indicates that \texttt{U} is a two-dimensional array, with a halo region
whose size is obtained by the compiler from the explicit halo size of the actual array argument
provided at the calling site of the function.  In LOPe, there is no need (in this instance) for a
repetitive \texttt{dimension(:,:)} specification, as it is inferred from the \texttt{HALO}
specification.

\textbf{Concurrent functions.}
The second keyword employed by LOPe is \texttt{concurrent}.  This keyword already exists in the form
of a \texttt{do} \texttt{concurrent} loop construct.  By using this construct, the programmer
asserts that specific iterations of the loop body may be executed by the compiler in \emph{any
  order;} even \emph{concurrently.}
%%LOPe allows the \texttt{concurrent} attribute to be attached to a function declaration so that the function can be called from within a concurrent loop without fear of race conditions.
For example, the following,
\begin{verbatim}
  pure CONCURRENT subroutine Laplace(U)
     real, HALO(:,:) :: U
     U(0,0) =                 U(0,+1)               &
              +  U(-1,0)  - 3*U(0, 0)  +  U(+1,0)   &
                          +   U(0,-1)
  end subroutine Laplace
\end{verbatim}
defines a function that can be used as part of an iterative solution of Laplace's equation in two
dimensions.  LOPe allows a function with the \texttt{pure} and \texttt{concurrent} attributes to be
called from within a \texttt{do} \texttt{concurrent} loop body.  One should imagine that this
function is called \emph{for each} \texttt{(i,j)} index of the interior of the array \texttt{U}.
(An example of this usage will be given later in the text.)  Note that this usage introduces a race
condition as new values of elements of \texttt{U} are created on the left-hand side of the
assignment statement that may use \emph{new or old} values of \texttt{U} on the right-hand side.
LOPe requires the compiler to guarantee that race conditions won't occur by using, e.g.,
double-buffering techniques as needed.

\textbf{LOPe index notation.}
In the \texttt{Laplace} example shown above, the \texttt{U(0,0)} array element is the \emph{local}
array element and only the local element may be modified.  This zero-based indexing for the
local-array element is different than conventional Fortran, where by default, array indices start at
1.  For LOPe, this default convention is modified; indices for arrays declared with the
\texttt{HALO} attribute (within the context of a \texttt{CONCURRENT} procedure) range (in one
dimension) from \texttt{-M} to \texttt{+N} where the halo attribute has been declared as
\texttt{M:*:N} by the originating type declaration.  The other array elements (in the halo region of
the example) are \texttt{U(-1,0)} and \texttt{U(+1,0)} (left and right of local, respectively) and
\texttt{U(0,-1)} and \texttt{U(0,+1)} (below and above of local, respectively).  The geometric
positioning of the array elements can be seen by examining the arrangement of the expressions on the
right-hand side of the \texttt{Laplace} example.