### 2015-01-31 papers/LOPe/programming-model.tex

\section{Programming Model}

The LOPe programming model (for Locally-Oriented Programming
extensions) restricts the programmer to a local view of the index
space of an array.  Within a LOPe function, only a single array
element (called the local element) is mutable.  In addition, a small
halo region surrounding the local element is visible to the
programmer, but this region is immutable.  Restricting the programmer
to a local index space serves to reduce complexity by separating all
data- and task-decomposition concerns from the implementation of the
element-level array calculations.

%%This reduction in complexity reduces programming errors.  While developing the convolution example described later in the paper we made an indexing error in applying the 2D stencil loops in the standard serial Fortran test implementation.  This error required over 3 hours of programming time to repair.  First the error had to be isolated to the function implementing the convolution (it was first thought to be in the complicated tiff image output routine as the convolution code was ``thought'' to be too simple to wrong.  Then the index error has to be understood.  As will be seen, LOPe makes it more difficult to make these errors as Fortran array intrics can be used.  In some instance, the restricted semantics of LOPe allows the compiler to catch errors (e.g., some errors involving race conditions).

LOPe is a domain specific language (DSL) implemented as a small extension
to standard Fortran 2008\todo{What version of Fortran does it extend?}.
Fortran was chosen as the base language for LOPe, although in
principle, the same techniques could be applied to language such as C
or C++ (in conjuction with some form of array extensions such as those
proposed by Intel).  Fortran was chosen because it already provides a
rich array-based syntax.  Readers who are unfamiliar with Fortran
syntax may wish to consult Appendix A, for a brief description of
Fortran notation.

\subsection{Related work}

LOPe builds upon prior work studying how to map Fortran to accelerator
programming models like OpenCL.  In the ForOpenCL project~\cite{foropencl}
we exploited Fortran pure and elemental functions to express the core
data parallel kernels of an OpenCL-based program.  In practice, array
calculations for a given index $i,j$ will require read-only access to
a local neighborhood of size $[M,N]$ around $i,j$.  The LOPe work introduces
a mechanism for representing these neighborhoods as array declaration
type annotations.

This work is based on concepts explored in the ZPL programming
language~\cite{zpl1,zpl2} in which the programmer can define regions and 
operators that are applied over the index sets corresponding to the regions.
For example, the following ZPL code implements the same stencil as the
LOPe code in Section~\ref{sec:where-is-jacobi-example}:

\begin{verbatim}
ZPL jacobi example here.
\end{verbatim}

This approach as demonstrated by ZPL is quite powerful for compilation
purposes since it provides a clean decoupling of the operators applied
over an array from the decomposition of that array over a potentially
complex distributed memory hierarchy.

\subsection{LOPe Syntax Extensions}

There are only a few syntax additions required for a LOPe program.
These additions include syntax for describing halo regions and
concurrent procedures.  In all of the code examples that follow,
syntax additions are highlighted by the usage of capitalization for
keywords that are either new or that acquire new usage.

\textbf{Halo regions.}

The principle semantic element of LOPe is the concept of a halo
region.  A halo is an ``artificial'' or ``virtual'' region surrounding
an array that contains boundary-value information.  Halo (also called
ghost-cell) regions are commonly employed to unify array indexing
schemes in the vicinity of an array boundary, so that an array may be
referenced using indices that fall ``outside'' of the logical domain
of the array.  In LOPe, the halo region is given explicit syntax so
that the compiler can exploit this information for purposes of memory
allocation, data replication and thread synchronization.  For example,
a halo region can be declared with a statement of the form,

\begin{verbatim}
  real, allocatable, dimension(:), HALO(1:*:1) :: A
\end{verbatim}

This statement indicates that \texttt{A} is a rank one array, will be
allocated later (when it will be provided with a specific size), and
has a halo region of one element surrounding the array on either side.
Notation such as \texttt{M:*:N} is to be read as describing a halo
region of \texttt{M} elements to the left, \texttt{N} elements to the
right, and with an arbitrary number of elements in the middle (denoted
by the $*$ symbol).  When used to describe a formal parameter of a
function, the halo size may be ``assumed'' (denoted by a ``:'' symbol
substituted in place of an explicit size) as is the following,

\begin{verbatim}
  real, HALO(:,:) :: U
\end{verbatim}

which in this instance indicates that \texttt{U} is a two-dimensional
array, with a halo region whose size is obtained by the compiler from
the explicit halo size of the actual array argument provided at the
calling site of the function.  In LOPe, there is no need (in this
instance) for a repetitive \texttt{dimension(:,:)} specification, as
it is inferred from the \texttt{HALO} specification.\todo{This paragraph is a bit weird to read.}

\textbf{Concurrent functions.}
The second keyword employed by LOPe is \texttt{concurrent}.  This keyword already exists in the form
of a \texttt{do} \texttt{concurrent} loop construct.  By using this construct, the programmer
asserts that specific iterations of the loop body may be executed by the compiler in \emph{any
  order;} even \emph{concurrently.}
%%LOPe allows the \texttt{concurrent} attribute to be attached to a function declaration so that the function can be called from within a concurrent loop without fear of race conditions.
For example, the following,
\begin{verbatim}
  pure CONCURRENT subroutine Laplace(U)
     real, HALO(:,:) :: U
     U(0,0) =                 U(0,+1)               &
              +  U(-1,0)  - 3*U(0, 0)  +  U(+1,0)   &
                          +   U(0,-1)
  end subroutine Laplace
\end{verbatim}

defines a function that can be used as part of an iterative solution
of Laplace's equation in two dimensions.  LOPe allows a function with
the \texttt{pure} and \texttt{concurrent} attributes to be called from
within a \texttt{do} \texttt{concurrent} loop body.\todo{Worth pointing out that pure implies no side effects, concurrent implies no dependencies between iterations?}  One should
imagine that this function is called \emph{for each} \texttt{(i,j)}
index of the interior of the array \texttt{U}.  (An example of this
usage will be given later in the text.)  Note that this usage
introduces a race condition as new values of elements of \texttt{U}
are created on the left-hand side of the assignment statement that may
use \emph{new or old} values of \texttt{U} on the right-hand side.
LOPe requires the compiler to guarantee that race conditions won't
occur by using, e.g., double-buffering techniques as needed.

\textbf{LOPe index notation.}

In the \texttt{Laplace} example shown above, the \texttt{U(0,0)} array
element is the \emph{local} array element and only the local element
may be modified.  This zero-based indexing for the local-array element
is different than conventional Fortran, where by default, array
indices start at 1.  The use of zero-based indexing gives a clean
symmetry for indices on either side of the central element at zero.
For LOPe, this default convention is modified; indices for arrays
declared with the \texttt{HALO} attribute (within the context of a
\texttt{CONCURRENT} procedure) range (in one dimension) from
\texttt{-M} to \texttt{+N} where the halo attribute has been declared
as \texttt{M:*:N} by the originating type declaration.  The other
array elements (in the halo region of the example) are
\texttt{U(-1,0)} and \texttt{U(+1,0)} (left and right of local,
respectively) and \texttt{U(0,-1)} and \texttt{U(0,+1)} (below and
above of local, respectively).  The geometric positioning of the array
elements can be seen by examining the arrangement of the expressions
on the right-hand side of the \texttt{Laplace} example.