### 2015-02-04 papers/LOPe/intro.tex
Many changes to serve as a candidate for an introduction.

\section{Introduction}

Historically it has been hard to create parallel programs.  The responsibility (and difficulty) of
creating \emph{correct} parallel programs can be viewed as being spread between the programmer and
the compiler.  Ideally we would like to have parallel languages that make it easy for the
programmer to express correct parallel programs (and conversely it should be difficult to express
incorrect parallel programs).  Unfortunately, many current languages and standards place all of the
responsibility on the user.  The best example of this are programs written using MPI (Message
Passing Interface), where the programmer expresses all parallelism in terms of calls to library
routines and the serial C or Fortran compiler knows nothing about the parallel semantics of the
combined language plus MPI library.

For the purpose of discussion, we define a degree of difficulty term called the joint Parallel
Complexity Product (PCP),

\begin{verbatim}
    PCP = code_complexity X compiler_complexity
\end{verbatim}
and suggest that PCP is roughly a constant over time, say the level of complexity of writing an MPI
program today, PCP\_MPI.

Of course one would hope that over time code complexity goes down as better languages allow
compilers to take on a greater share of the parallel complexity burden.  This is somewhat true of
Unified Parallel C (UPC) and Coarray Fortran.  With the PGAS extensions, C and Fortran compilers
are now aware of parallelism and now generate message passing code that previously had been handled
by the MPI library.  In some instances the compiler is able to perform optimizations that it is not
able to do with a library-based scheme like MPI \cite{preissl:2011:SC}.

%%% Would like to say something here about HPC as a language based solution to reducing code complexity

However, in large part these languages are largely syntatic sugar for message passing and do not
provide a substatial decrease in code complexity when application performance is a
goal\cite{hasert:2011:EuroMPI}.  While skilled programmers in the HPC community have become
accustomed to the level of complexity of an MPI program,
%%and yet are welcome to considering the simplifications of UPC and Coarray Fortran.
the problem for programmers is that hardware is changing in ways that increase the level of on-chip
parallelism.  Indeed, the current generation of machines could be the last to provide homogeneous
multi-core parallelism\cite{JORSjors.aw}.  For ultimate performance, programmers must envision
coding for heterogeneous architectures 
%%hyper-parallelism
with huge new levels of on-node parallelism, at the same time they account for off-node
parallelism.  Thus PCP has suddenly increased with PCP $=$ PCP\_MPI $+$ PCP\_Multi\_Core $>>$
PCP\_MPI.  Since languages have not evolved that allow the compiler to take up this increase, the
complexity for a programmer has necessarily dramatically increased.  Unfortunately, considering the
complexity of large multiphysics code frameworks, it could take several person years to retarget an
application for \emph{each} new heterogeneous platform\cite{DBLP:journals/corr/DubeyS13}.

A reasonable solution given todays language limitations is to use MPI for distributed memory
parallelism \emph{plus} OpenMP for on-chip parallelism.  This hybrid approach is taken by two
successful, large-scale AMR multiphysics code frameworks, FLASH and Chombo; currently however,
neither of these codes currently support hardware accelerators\cite{DBLP:journals/corr/DubeyS13}.
%%This is the solution proposed by Cray and PGI \cite{BOF_SC10} (currently Chapel does not target GPUs\cite{Brad?}).
Other choices for expressing on-chip parallelism are OpenCL\cite{OPENCL} and NVIDIA's CUDA.
Unfortunately, achieving high performance using either of these two languages can be a daunting
challenge and there are no guarantees that either language is suitable for future hardware
generations.

In this paper we examine a language-based paradigm that allows the compiler to take on a larger
portion of the PCP burden.  Anyone programming in OpenCL or CUDA is aware that explicit loop
structures over array elements in a serial program are removed and replaced by a kernel program
that is run over all of the elements of the input arrays.  We propose Locally Orientated
Programming extensions (LOPe) to the Fortran and potentially C languages that formally adopt this
programming model.  The LOPe programming model explicitly separates numerical complexity from
parallel complexity --- leaving much of the staging and assembling of data to the compiler --- and
is applicable to the solution of stencil-based algorithms that provide data access patterns that
are regular and spatially local.