### 2015-02-05 papers/LOPe/transformations.tex
Added very little text.

\section{LOPe Implementation}

This section describes how LOPe extensions to Fortran have been implemented.  It also describes how
source-to-source transformations have been implemented via rewrite rules and strategies in
StrategoXT.

\subsection{Halo Transfer Performance}

Amortization of halo transfer and kernel