### 2015-02-07 papers/LOPe/local.tex
Paper lenght shortening begun.

%%%%%%%%%%%%%%%%%%%%%%% file typeinst.tex %%%%%%%%%%%%%%%%%%%%%%%%%
%
% This is the LaTeX source for the instructions to authors using
% the LaTeX document class 'llncs.cls' for contributions to
% the Lecture Notes in Computer Sciences series.
% http://www.springer.com/lncs       Springer Heidelberg 2006/05/04
%
% It may be used as a template for your own input - copy it
% to a new file with a new name and use it as the basis
% for your article.
%
% NB: the document class 'llncs' has its own and detailed documentation, see
% ftp://ftp.springer.de/data/pubftp/pub/tex/latex/llncs/latex2e/llncsdoc.pdf
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass[runningheads,a4paper]{llncs}

\usepackage{amssymb}
\setcounter{tocdepth}{3}
\usepackage{graphicx}
\usepackage{todonotes}
\usepackage{url}
%%\urldef{\mailsa}\path|rasmus@uoregon.edu, mjsottile@gmail.com,|
\urldef{\mailsa}\path|rasmus@uoregon.edu|
\urldef{\mailsb}\path|dangle@ucar.edu, soren.rasmussen@aggiemail.usu.edu|
\newcommand{\keywords}[1]{\par\addvspace\baselineskip
\noindent\keywordname\enspace\ignorespaces#1}

\usepackage{cite}

\begin{document}

\mainmatter  % start of an individual contribution

% first the title is needed
\title{Locally-Oriented Programming:\\ A Simple Programming Model for Stencil-Based Computations
       on Multi-Level Distributed Memory Architectures}

% a short form should be given in case it is too long for the running head
\titlerunning{Locally-Oriented Programming}

\author{Craig Rasmussen
\and Matthew Sottile\and Daniel Nagle\and Soren Rasmussen}

\authorrunning{Locally-Oriented Programming}
% (feature abused for this document to repeat the title also on left hand pages)

% the affiliations are given next; don't give your e-mail address
% unless you accept that it will be published
%%\institute{441 McKenzie Hall,\\
%%5246 University of Oregon, Eugene, Oregon, USA\\
%%\mailsa\\
%%\mailsb\\
%%\url{https://openfortran.uoregon.edu}}

\institute{University of Oregon, Eugene, Oregon, USA\\
\mailsa
}

%
% NB: a more complex sample for affiliations and the mapping to the
% corresponding authors can be found in the file "llncs.dem"
% (search for the string "\mainmatter" where a contribution starts).
% "llncs.dem" accompanies the document class "llncs.cls".
%

%\toctitle{Lecture Notes in Computer Science}
%\tocauthor{Authors' Instructions}
\maketitle

\begin{abstract}

Emerging hybrid accelerator architectures for high performance computing are often suited for the
use of a data-parallel programming model.  Unfortunately, programmers of these
architectures face a steep learning curve that frequently requires learning a new language (e.g.,
OpenCL). Furthermore, the distributed (and
frequently multi-level) nature of the memory organization of clusters of these machines
provides an additional level of complexity.  This paper presents preliminary work
examining how programming with a local orientation can be employed to provide simpler access to
accelerator architectures.  A locally-oriented programming model is especially useful for the
solution of algorithms requiring the application of a stencil or convolution kernel.  In this programming
model, a programmer codes the algorithm by modifying \emph{only a single array element} (called the
local element), but has read-only access to a small sub-array surrounding the local element.
We demonstrate how a locally-oriented programming model can be adopted as
a language extension using source-to-source program transformations.

\keywords{domain specific language, stencil compiler, distributed memory parallelism}
\end{abstract}

%\input{motivation}
\input{intro}
\input{programming-model}
\input{coarray-extensions}
\input{transformations}
\input{conclusions}

\bibliographystyle{plain}
\bibliography{local}{}

%\begin{thebibliography}{4}
%\end{thebibliography}

\subsubsection*{Acknowledgments.}
This work was supported in part by the Department of Energy Office of Science, Advanced Scientific
Computing Research.  The authors would also like to thank Robert Robey and Wayne Weseloh at Los
Alamos National Laborabory for several stimulating conversations with respect to programming models.

\input{appendix-a}

\end{document}