### 2016-01-07 papers/CAFe/HIPS-paper/intro.tex

\section{Introduction}

A goal of high-level programming languages should be to \emph{allow} a programmer to
develop software for a specific hardware architecture, without actually \emph{exposing} hardware
features.  In fact, allowing low-level hardware details to be controlled by the language is actually
a barrier to software performance portability.  The path forward for high-level
programming languages should be to allow a programmer to provide more symantic information
about the intent of the programmer, thus allowing the compiler more freedom to retarget
applications to new architectures.

This becomes critical as the scientific community tries to achieve exascale performance.
Energy is a major concern, thus, heterogenous archtectures with many simple, low-energy cores.

There is a plethera of options available to an HPC programmer beyond just using a serial
language with MPI.  New parallel languages are under development, for example, Chapel and
the PGAS languages including Coarray Fortran (CAF).  Options for heterogeneous computing
include usage of OpenMP, OpenACC, CUDA, CUDA Fortran, and OpenCL for programming attached
accelerator devices.

A reasonable solution given today's language limitations is to use MPI for distributed memory
parallelism \emph{plus} OpenMP for on-chip parallelism.  This hybrid approach is taken by two
successful, large-scale AMR multiphysics code frameworks, FLASH and Chombo; currently however,
neither of these codes support hardware accelerators\cite{DBLP:journals/corr/DubeyS13}.
%%This is the solution proposed by Cray and PGI \cite{BOF_SC10} (currently Chapel does not target GPUs\cite{Brad?}).
Other choices for expressing on-chip parallelism are OpenCL~\cite{opencl11} and NVIDIA's CUDA.
Unfortunately, achieving high performance using either of these two languages can be a daunting
challenge and there are no guarantees that either language is suitable for future hardware
generations.

OpenACC directives and features:
\begin{verbatim}
● acc_copy(in|out)( ptr, bytes )
● acc_create( ptr, bytes )
● acc_delete( ptr, bytes )
● acc_is_present( ptr, bytes )
● acc_update_(device|local)( ptr, bytes )
● acc_deviceptr( ptr )
● acc_hostptr( devptr )
● acc_[un]map_data( devptr, hostptr, bytes )
● acc_memcpy_(to|from)_device
\end{verbatim}

With the addition of coarrays in the 2008 standard, Fortran is now a \emph{parallel}
language.  Originally called Coarray Fortran (CAF) by its
authors\cite{Numrich:1998:CFP:289918.289920}, parallism in Fortran is similar to MPI in
that all cooperating CAF processes (called images) execute a Single Program but with
different (Multiple) Data (SPMD).  Key concepts in CAF are parallel execution, distributed
memory allocation, and remote memory transfer.  Parallel execution and memory allocation
are straightforward as each program is replicated $NP$ times with each program image
allocating its own memory locally.  Special syntax is provided to indicate which arrays
are visible between distributed memory processes and remote memory transfer is indicated
with new array syntax using square brackets, for example,
\begin{verbatim}
   U(1) = U(1)[3]
\end{verbatim}
copies memory from the first element of \texttt{U} residing on image number 3 into
the corresponding element of the \emph{local} array \texttt{U}.

Since each CAF process executes the same program, there are no good avenues for exploiting
task-based parallelism.  Neither is there a mechanism for dynamic coarray memory
allocation, nor a mechanism for the remote execution of functions.  In CAF, like MPI, the
programmer often combines distributed memory processes running the same program image with a
shared memory threaded programming model using OpenMP (or OpenACC).

In this paper we examine...

Topics highlighted in this section are: 1. Distributed memory array allocation;
2. Explicit memory placement; 3. Remote memory transfer; and 4. Remote execution.  This
description is within the context of extensions to Fortran; as shorthand, these extensions
are referred to as CAFe, for Coarray Fortran extensions.  CAFe is complementary to
previous work extending coarray Fortran\cite{mellor-crummey:2009:caf2,jin:2011:caf2}.

Historically it has been hard to create parallel programs.  The
responsibility (and difficulty) of creating \emph{correct} parallel
programs can be viewed as being spread between the programmer and the
compiler.  Ideally we would like to have parallel languages that make
it easy for the programmer to express correct parallel programs --- and
conversely it should be difficult to express incorrect parallel
programs.  Unfortunately, many current languages and standards place
all of the responsibility on the user.  The best example of this are
programs written using MPI (Message Passing Interface), where the
programmer expresses all parallelism in terms of calls to library
routines and the serial C or Fortran compiler knows nothing about the
parallel semantics of the combined language plus MPI library.

Of course one would hope that over time code complexity goes down as
better languages allow compilers to take on a greater share of the
parallel complexity burden.  Multithreaded code written using OpenMP
can be significantly simpler than the corresponding code in which the
programmer explicitly manages a pool of threads themselves.  This is
somewhat true for distributed memory parallelism using Unified
Parallel C (UPC) and Coarray Fortran.  With these PGAS
(Partitioned Global Address Space) extensions, C and
Fortran compilers are now aware of parallelism and now generate
message passing code that previously had been handled by the MPI
library.  In some instances the compiler is able to perform
optimizations that it is not able to do with a library-based scheme
like MPI \cite{preissl:2011:SC}.

%%% Would like to say something here about HPC as a language based solution to reducing code complexity

However, in large part these languages are mostly syntactic sugar for
message passing and do not provide a substantial decrease in code
complexity when application performance is a
goal\cite{hasert:2011:EuroMPI}.  While skilled programmers in the HPC
community have become accustomed to the level of complexity of an MPI
program, the problem for programmers is that hardware is changing in
ways that increase the level of on-chip parallelism.  Indeed, the
current generation of machines could be the last to provide
homogeneous multi-core parallelism\cite{JORSjors.aw}.  For ultimate
performance, programmers must envision coding for heterogeneous
architectures with huge new levels of on-node parallelism, at the same
time they account for off-node parallelism.
Since languages have not evolved that allow the compiler to take up
this increase, the complexity for a programmer has necessarily
dramatically increased.  Unfortunately, considering the complexity of
large multiphysics code frameworks, it could take several person years
to retarget an application for \emph{each} new heterogeneous
platform\cite{DBLP:journals/corr/DubeyS13}.

In this paper we examine a language-based paradigm that allows the
compiler to take on a larger portion of the code complexity burden.
Anyone programming in OpenCL is aware that explicit loop
structures over array elements in a serial program are removed and
replaced by a kernel program that is run over all of the elements of
the input arrays.  We propose \emph{Locally Orientated Programming
extensions} (LOPe) to the Fortran and potentially C languages that
formally adopt this programming model.  The LOPe programming model
explicitly separates numerical complexity from parallel complexity ---
leaving much of the staging and assembling of data to the compiler ---
and is applicable to the solution of stencil-based algorithms that
provide data access patterns that are regular and spatially local.