### 2016-01-09 papers/CAFe/HIPS-paper/performance.tex

\section{Performance Measurements}

This section shows and compares performance measurements for: 1. Memory bandwidth usage on GPU; 2. Computaional time on GPU; 3. Halo Communication time (total plus breakdown to and from device and MPI); 4. Total time scaling on ACISS; 5. Comparison with OpenACC (hopefully).