### 2016-01-19 papers/CAFe/HIPS-paper/performance.tex
Added performance table and text.

\section{Performance Measurements}

It is important to note that the emphasis of this work is to explore and explain the new
syntax and execution semantics of a CAFe application.  A thorough examination of potential
performance gains (if any) using CAFe for parallelization of code is beyond the scope of
this paper.  The primary purpose of CAFe is to combine the parallel features of Fortran coarrays
--- executing a \emph{single} program --- with concurrent execution of separate tasks executing on
potentially heterogeneous hardware.

\begin{comment}
However, the relative performance of computation on a cluster of GPUs compared with the
necessary communication of halo information is of interest, especially considering that a
complete exchance of halo data involves communication between multiple coarray images
\emph{and} between each individual host image and its subimage (an attached OpenCL
device).
\end{comment}

However the relative performance of computation on the interior of a three-dimensional
grid, performed by the GPU, compared with the time required to compute on boundary planes by the
CPU is of interest.  Table 1 shows average execution time for a relaxation step on the GPU
(column 2) and on the CPU (column 3) in milliseconds, where $N$ (column 1) is the number of
cells in one dimension of a 3D cube; time for the exchange of halo information
between the GPU and the CPU for each iteration is also shown in column 4.

\begin{table}[]
\centering
\caption{Size of cube and time in milliseconds}
\label{table1}
\begin{tabular}{rrrrr}
$N$   &   GPU	 &  CPU         &   Comm    \\
      &   	 &              &           \\
16    &  0.02    &   0.01	&   0.07    \\
32    &  0.02    &   0.03       &   0.10    \\
64    &  0.04    &   0.12       &   0.27    \\
128   &  0.42    &   0.47	&   0.81    \\
256   &  4.59    &   1.97	&   2.99    \\
512   & 43.08    &   12.51	&  16.49    \\
\end{tabular}
\end{table}

Note that all three times in Table 1 are roughly equivalent for $N=128$.  Also note that
the communication and the computational times on the CPU are roughly equivalent and scale
the same with increasing $N$.  This is not surprising as both involve only the surface
elements of the cube.  However, computational time on the GPU grows much faster, O($N^3$),
as it is computing on the interior of the cube.

\begin{comment}
The current implementation does \emph{not} take advantage of optimization strategies such
as prefetching of array tiles (including halos) into OpenCL local memory.  Neither does it
take advantage of the potential of CAFe to overlap communication with computation (for
example, computing on ...
\end{comment}

\begin{comment}
Since many scientific codes are dominated by memory performance, including and especially
stencil algorithms as they typically only involve a computation on a small locally central
array element and a small overlapping halo region.  Stencil operations frequently do not
contain enough floating point operations per memory load to allow for floating point
performance to operate at peak (though this is entirely application and domain specific).
Thus we illustrate the \emph{potential} for performance by noting the latency and
throughput performance of an attached GPU in conjunction with MPI distributed memory
performance associated with halo transfer in Table 1.
\end{comment}

\begin{comment}
The results in Table 1 indicate that the primary bottleneck in using accelerators attached
to the TODO bus using OpenCL is likely to be the latency in transferring memory to and
from the device for distributed memory clusters of only a few nodes exchanging halo data
using MPI.
\end{comment}